# Stock basics {#sec-stock-basics}

```{r}
#| include: false
source(here::here("_common.R"))

by_type <- stock_by_type(shop)
on_hand <- stock_on_hand(shop) |>
  left_join(shop$products, by = "sku")
temple_jhumka <- filter(on_hand, sku == "EAR-0002")
antique_jhumka <- filter(on_hand, sku == "EAR-0001")
total_at_cost <- sum(by_type$value_at_cost)
total_at_price <- sum(by_type$value_at_price)
sets_share <- by_type$value_at_cost[by_type$type == "Sets"] / total_at_cost
# As the Overview writes it: pairs, then sets, then pieces.
units_line <- by_type |>
  summarise(units = sum(units), .by = unit) |>
  arrange(factor(unit, levels = c("pair", "set", "piece"))) |>
  mutate(words = describe_units(units, unit)) |>
  pull(words) |>
  paste(collapse = " · ")
all_units <- sum(by_type$units)
```

::: {.audience}
For everyone
:::

::: {.learn}
- What a product is, and what its SKU is for.
- Why earrings are counted in pairs and sets as one.
- What cost, list price and stock value mean.
:::

## The idea

Your [**stock**](glossary.qmd#stock) is the jewellery the shop owns and means to sell. To look after
it, you need to know three things about every piece: what it is, how many of
it there are, and what it's worth. This chapter covers the first and the
last. How many there are is the subject of the next two chapters.

### Products and SKUs

Each design the shop sells is a [**product**](glossary.qmd#product). A product is one design, in one
metal and one weight, so that every piece of it is alike. The Temple Lakshmi
jhumka in 22K gold weighing 14 g is one product. The antique jhumka in 22K
weighing 8 g is another, even though both are jhumkas.

Each product has a short code called its [**SKU**](glossary.qmd#sku), which stands for *stock
keeping unit*. The Temple Lakshmi jhumka's SKU is EAR-0002. A SKU belongs to
one product only. You can change it with Edit, but it is best to keep it the
same, because labels and files use it. Jholok treats small and capital
letters as the same, so `ear-0002` is the same SKU as `EAR-0002`. When you add
a product, Jholok suggests a SKU from its type (EAR for earrings, RIN for
rings), or you can use your own codes.

How many of a product the shop holds is its [**stock on hand**](glossary.qmd#on-hand).

### Types and units

Products are grouped into [**types**](glossary.qmd#type): in the demo shop these are Earrings,
Rings, Necklaces, Bangles, Sets, Coins and Other. A type decides the [**unit**](glossary.qmd#unit)
its products are counted in:

- **Earrings are counted in [pairs](glossary.qmd#pair).** A pair is sold together, so it is
  counted together. A [single earring](glossary.qmd#odd-single) found on its own is not half a pair,
  so it is never rounded into one: count the whole pairs, and add the single
  with **Singles** in the count sheet. It shows beside the pairs ("5 pairs +
  1 single"), the approver sees it in Review, and it never changes stock.
- **[Sets](glossary.qmd#set) are counted as one.** A necklace with its matching earrings is one
  set, and counts as 1, because it is sold as one.
- **Everything else is counted in [pieces](glossary.qmd#piece)**: a ring, a chain, a coin.

Because the units are different, Jholok keeps them apart when it says what
is in stock: "`r units_line`", not one total. Adding a pair of studs to a
bridal set gives a number that means little.

Jholok never adds them into one count of [**items**](glossary.qmd#item), even
for sales. Where several kinds sold, it lists each, such as "18 pairs and 13
pieces". Where a rupee value says enough, as in "Last week sold ₹7.79 L", it
gives the value alone.

### Cost, list price and value

Each product can have two amounts:

- its [**cost**](glossary.qmd#cost): what the shop paid for one, with the gold, the [making charge](glossary.qmd#making-charge)
  and any stones included;
- its [**list price**](glossary.qmd#list-price): what the shop asks a customer for one (the product page
  calls it Price).

[**Stock value**](glossary.qmd#stock-value) is what the stock is worth: each product's stock on hand
times its cost, added up. Value at cost is what the shop has paid for the
stock it holds. Value at list price is what it would take if everything sold
at the list price. The gap between the two is the shop's margin.

Large amounts are written in [lakh](glossary.qmd#lakh) and [crore](glossary.qmd#crore), as in the app. One lakh (L) is
₹1,00,000, and one crore (Cr) is 100 lakh, or ₹1,00,00,000.

## In the app

The **Catalogue** tab lists the products, grouped by type.

```{r}
screenshot("catalogue.png")
```

Tap a product to open its page. The Temple Lakshmi jhumka has
`r describe_units(temple_jhumka$on_hand, "pair")` on hand. Its value,
`r rupees_short(temple_jhumka$on_hand * temple_jhumka$cost)` at cost, is
`r temple_jhumka$on_hand` × `r rupees(temple_jhumka$cost)`. [**Cover**](glossary.qmd#cover) says how
long the stock lasts at the pace it has sold recently.

```{r}
screenshot("product-detail.png")
```

The **Overview** tab shows the whole shop: its stock value at cost,
`r rupees_short(total_at_cost)`, and at list price,
`r rupees_short(total_at_price)`, and what is in stock, by unit.

```{r}
screenshot("overview.png")
```

## Why it works this way

- **Stock is kept in its units** so that every total means something. "20
  sets" tells you something; "`r all_units` items" doesn't.
- **A value always says what it is based on, and when.** Value at cost and
  value at list price differ by the shop's margin, and gold prices move. So
  Jholok writes "`r rupees_short(total_at_cost)` at today's cost, `r date_words(screens_taken_on)`", with
  its date, never just a figure.
- **Nothing is left out quietly.** One product in the demo shop has no cost,
  so the Overview says "Leaves out 1 product with no cost" rather than
  counting it as ₹0.

## How it works

For each product, Jholok multiplies its stock on hand by its cost (or its
list price) and adds the results. Stock on hand itself isn't stored anywhere:
Jholok works it out from the ledger, as the next chapter explains. If a
product's stock has gone below zero (a mistake the next chapter also covers),
it counts as none in the value, because no shelf holds minus one pendant.

## Try it

This is Jholok Jewellers' stock at the end of the demo's story, by type.

```{r}
#| label: tbl-stock-by-type
#| tbl-cap: >-
#|   Jholok Jewellers' stock by type. Other leaves out 1 product with no cost,
#|   and Coins 1 with no list price.
by_type |>
  transmute(
    Type = type,
    `In stock` = units_words,
    `At cost` = rupees_short(value_at_cost),
    `At list price` = rupees_short(value_at_price)
  ) |>
  table_guide() |>
  gt::cols_align(align = "right", columns = -Type)
```

The shop holds only `r by_type$units[by_type$type == "Sets"]` sets, but they
hold `r percent_words(sets_share)` of its value at cost, because each set is
heavy and costly.

**Question.** The antique jhumka (EAR-0001) costs
`r rupees(antique_jhumka$cost)` a pair, and the shop has
`r describe_units(antique_jhumka$on_hand, "pair")`. What are they worth at
cost?

::: {.callout-tip collapse="true"}
## Answer

`r antique_jhumka$on_hand` × `r rupees(antique_jhumka$cost)` =
`r rupees(antique_jhumka$on_hand * antique_jhumka$cost)`, which the app writes
as `r rupees_short(antique_jhumka$on_hand * antique_jhumka$cost)`.
:::

::: {.remember}
- One design is one product, with its own SKU.
- Earrings are counted in pairs, sets as one and the rest in pieces, and
  stock is shown in those units, not as one total.
- Stock value is stock on hand times cost (or list price). It always says
  which, and from when.
:::
