# The ledger: why stock is a sum {#sec-ledger}

```{r}
#| include: false
source(here::here("_common.R"))

jhumka_moves <- stock_history(shop, "EAR-0001")
jhumka_out <- filter(shop$out_records, sku == "EAR-0001")
jhumka_in_shop <- stock_in_shop(shop) |>
  filter(sku == "EAR-0001") |>
  pull(in_shop)

# The ledger shows a send-out as one row, the shop's side, with where it went.
jhumka_rows <- jhumka_moves |>
  filter(!(kind == "sent_out" & is.na(place))) |>
  mutate(
    shown = if_else(
      kind == "sent_out",
      paste(quantity, "to", place),
      sub("-", minus_sign, sprintf("%+d", quantity), fixed = TRUE)
    )
  )
jhumka_by_kind <- jhumka_moves |>
  summarise(change = sum(quantity), .by = kind)
change_of <- \(kind_name) {
  sum(jhumka_by_kind$change[jhumka_by_kind$kind == kind_name])
}

pendant <- stock_history(shop, "NEC-0003")
pendant_last_sale <- pendant |>
  filter(kind == "sold", on_hand == 0) |>
  slice_tail(n = 1)
pendant_early_sale <- slice_tail(pendant, n = 1)

last_30_days <- shop$movements |>
  filter(date > shop$end_at - 30 * 86400) |>
  summarise(units = sum(abs(quantity)), .by = kind)
units_of <- \(kind_name) {
  sum(last_30_days$units[last_30_days$kind == kind_name])
}

signed <- \(n) sub("-", minus_sign, sprintf("%+d", n), fixed = TRUE)
```

::: {.audience}
For everyone
:::

::: {.learn}
- What the ledger is, and what a movement is.
- Why Jholok never lets anyone type in a stock number.
- What it means when stock goes below zero.
:::

## The idea

A [**ledger**](glossary.qmd#ledger) is a record of every change to the stock, oldest first. Each
change is a [**movement**](glossary.qmd#movement). A movement says which product, when, and what kind of
change it was. It also says how many: a plus number for stock coming in, a
minus number for stock going out. It has a note, and the name of who entered
it.

These are the kinds of movement you'll meet first:

| Movement | What happened | Stock |
|---|---|---|
| [Opening stock](glossary.qmd#opening-stock) | The product was added to Jholok with stock already in the shop | goes up |
| Received | A delivery came in from a supplier or the [karigar](glossary.qmd#karigar) | goes up |
| Sold | A customer bought it | goes down |
| Returned | A customer brought it back | goes up |
| [Adjustment](glossary.qmd#adjustment) | A correction outside a count, with a reason, such as a piece damaged and sent for melting | up or down |
| Count adjustment | A count found a difference and the approver accepted it (@sec-counting) | up or down |

There are a few more, for loading the shop's billing records and for stock
that leaves the shop for a while. Later chapters explain them.

**Stock on hand is the sum of all a product's movements.** It works like a
bank passbook: the balance is every deposit minus every withdrawal, and
nobody writes in a new balance by hand.

Here is the antique jhumka's year. It opened with
`r change_of("opening")` pairs, `r change_of("received")` more were received,
`r change_of("returned")` was returned and `r -change_of("sold")` were sold. A
count found one pair fewer than the ledger expected, and the approver
accepted it. That leaves
`r describe_units(sum(jhumka_moves$quantity), "pair")` on hand. Two of them
are out on [jangad](glossary.qmd#jangad) with `r jhumka_out$party`, so the product's page shows
`r jhumka_in_shop` in the shop and `r jhumka_out$units` off the shelf: still
the shop's, but not on the shelf (@sec-movements).

```{r}
#| fig-cap: >-
#|   The antique jhumka (EAR-0001) over the year: every kind of movement added
#|   together gives the stock on hand.
#| fig-alt: >-
#|   A bar chart that steps up for opening stock, deliveries and returns, steps
#|   down for sales and a count adjustment, and ends at the stock on hand.
#| fig-height: 3.4
diagram_ledger(shop, "EAR-0001")
```

## In the app

The **Ledger** tab lists every movement, newest first. At the top it sums up
the last 30 days. In the demo shop that is `r units_of("sold")` sold,
`r units_of("returned")` returned and `r units_of("received")` received.

```{r}
screenshot("ledger.png")
```

To record a movement by hand, open the product and tap **Receive**, **Sell**
or **Adjust**. The kinds are also at the top of the screen, where you choose
**Returned** for a customer's return. For Received, Sold or Returned, enter how
many. For Adjust, enter the stock you really have and a reason. Check the
date: it should be when the movement happened, which may be earlier than when
you enter it. Once People is set up (@sec-setting-up), **Entered by** shows the
owner; anyone else taps **Change** and enters their PIN, so the ledger shows
their name.

```{r}
screenshot("record-sale.png")
```

## Why it works this way

Suppose someone could simply change the antique jhumka's stock from 8 to 7.
The missing pair would vanish without a trace: nobody could tell whether it
was sold, lost, or a typing slip. With a ledger, every change has a date, a
kind, a note and a name. The question "where did it go?" always has an answer
to look up.

So a mistake is never erased. It is put right with another movement, and both
stay in the ledger. This unbroken history is called an [**audit trail**](glossary.qmd#audit-trail). It
lets the owner, an accountant or an insurer check the stock long after the
day.

## How it works

Jholok stores movements, never stock numbers. Whenever a screen shows stock,
Jholok adds up that product's movements up to that moment. The same sums give
the stock on any past date, which is how the Overview draws its charts.

### When stock goes below zero

A shelf can't hold fewer than nothing, so stock below zero always means a
movement is missing or wrong. The Lakshmi pendant (NEC-0003) sold its last
piece on `r date_words(pendant_last_sale$date)`. On
`r date_words(pendant_early_sale$date)` another sale was entered, from a
delivery of 6 that had arrived but hadn't been entered yet. The ledger now
says `r signed(sum(pendant$quantity))`.

```{r}
screenshot("product-below-zero.png")
```

Jholok doesn't hide this or round it to zero. It marks the product [**Below
zero**](glossary.qmd#below-zero), lists it under Needs attention on the Overview, and suggests what to
do: record the missing delivery, or count the product.

## Try it

These are the antique jhumka's last eight movements. The last column is the
running total of the stock the shop owns: each row's change added to the one
before. Sending pairs out on jangad doesn't change it, because they are still
the shop's.

```{r}
#| label: tbl-jhumka-moves
#| tbl-cap: >-
#|   The antique jhumka's last eight movements, with the stock after each.
jhumka_rows |>
  slice_tail(n = 8) |>
  transmute(
    Date = date_words(date, "%b %Y, %H:%M"),
    Movement = movement_name(kind),
    Pairs = shown,
    Note = note,
    `On hand` = on_hand
  ) |>
  table_guide()
```

**Question.** A customer returns a ring they bought yesterday. How do you
record it, and what happens to the ring's stock?

::: {.callout-tip collapse="true"}
## Answer

Open the ring, tap **Receive**, choose **Returned** at the top, and enter 1. Its stock goes up by
one. The sale stays in the ledger too, so the history shows both: sold, then
returned.
:::

::: {.remember}
- Every change to stock is a movement in the ledger, with a kind, a date, a
  note and a name.
- Stock on hand is the sum of the movements. Nobody types in a stock number.
- Mistakes are put right with another movement, never erased.
- Stock below zero means a movement is missing. Find it, or count the
  product.
:::
