# Counting and checking against the books {#sec-counting}

```{r}
#| include: false
source(here::here("_common.R"))

# Showcase 1, the demo's last completed count.
showcase <- shop$counts |>
  filter(location == "Showcase 1")
showcase_lines <- shop$count_lines |>
  filter(count_id == showcase$count_id) |>
  left_join(
    stock_on_hand(shop, showcase$started_at) |>
      rename(expected = on_hand),
    by = "sku"
  ) |>
  left_join(select(shop$products, sku, name, variant, cost), by = "sku") |>
  mutate(difference = counted - expected)
showcase_matched <- sum(showcase_lines$difference == 0)
showcase_share <- share(showcase_matched, nrow(showcase_lines))
showcase_short <- filter(showcase_lines, difference < 0)
showcase_over <- filter(showcase_lines, difference > 0)

# A count of the earrings and rings at the end of the story, with one pair
# really missing, one ring counted twice and a single earring found alone.
missing_pair <- tibble(sku = "EAR-0005", change = -1L)
counted_twice <- tibble(sku = "RIN-0001", change = 1L)
lone_earring <- tibble(sku = "EAR-0003", n = 1L)

first_count <- simulate_count(
  shop, shop$end_at,
  types = c("Earrings", "Rings"),
  shelf_changes = missing_pair,
  mistakes = counted_twice,
  odd_singles = lone_earring
) |>
  reconcile_count()
recount <- simulate_count(
  shop, shop$end_at,
  types = c("Earrings", "Rings"),
  shelf_changes = missing_pair,
  odd_singles = lone_earring
) |>
  reconcile_count()

line_of <- \(count, product_sku) filter(count, sku == product_sku)
studs <- line_of(first_count, "EAR-0005")
band <- line_of(first_count, "RIN-0001")
matched_first <- sum(first_count$result == "matched")
matched_recount <- sum(recount$result == "matched")
```

::: {.audience}
For everyone
:::

::: {.learn}
- What a count is for, and why counting is blind.
- What short and over mean, and the usual reasons for each.
- Why a shortage is counted again, and why someone who didn't count approves.
:::

## The idea

The ledger says what *should* be in the shop. A [**count**](glossary.qmd#count), sometimes called a
stock take, finds what *is* there. For each product, Jholok puts the two
numbers side by side:

- [**expected**](glossary.qmd#expected): what the ledger says should be there;
- **counted**: what the counters found.

Comparing them, and explaining every difference, is called
[**reconciliation**](glossary.qmd#reconciliation). A count can cover the whole shop, or just a few types or
one showcase.

### Short and over

When the two numbers agree, the product [**matched**](glossary.qmd#matched). When they don't:

- [**short**](glossary.qmd#short) means fewer were found than expected;
- [**over**](glossary.qmd#over) means more were found than expected.

A difference is a question, not yet an answer. These are the usual reasons:

| What happened | What the count shows |
|---|---|
| A piece was stolen, lost or damaged | short |
| A sale wasn't entered | short |
| A delivery wasn't entered | over |
| A return wasn't entered | over |
| A piece was put back in the wrong tray | short in one place, over in another |
| A counter missed a piece, or counted one twice | short or over |
| Something that isn't the shop's stock was in the tray, such as a customer's piece in for repair | over |

Stock that goes missing without being sold is called [**shrinkage**](glossary.qmd#shrinkage). A count
is how a shop finds out how much it has lost.

### Counting is blind

While a count is open, counters never see the expected numbers. People find
what they expect to find: a counter who knows there should be 9 chains tends
to stop at 9, or to look until a ninth turns up. A counter who doesn't know
just counts. That is called a [**blind count**](glossary.qmd#blind-count).

```{r}
screenshot(
  "counting.png",
  caption = "Counting the necklaces in Showcase 2: no stock, nothing expected."
)
```

### Two checks before stock changes

1. **A shortage is counted again first.** Review won't let anyone accept a
   shortage until the product is recounted, and the recount is blind too:
   the first count and the ledger stay hidden. Most shortages on a first
   count are counting slips, and a recount finds them before the stock is
   changed.
2. **Someone who didn't count approves.** The counter submits the count for
   review. Only an [**approver**](glossary.qmd#approver) who didn't count in it can open Review and
   complete it, signing with their PIN. This is called [**separation of
   duties**](glossary.qmd#separation-of-duties): the person who might have lost a piece isn't the one who decides
   it is lost.

```{r}
#| fig-cap: >-
#|   A count's life. Counting stops when it's submitted; the approver can send
#|   it back for more counting, with a reason.
#| fig-alt: >-
#|   Five steps in a row: Start, Count, Submit, Review and Complete, with a
#|   dashed arrow from Review back to Count labelled Reopen, with a reason.
#| fig-height: 2.4
diagram_count_life()
```

### What completing a count changes

Completing a count changes stock only for the differences the approver left
ticked. Each one becomes a [**count adjustment**](glossary.qmd#count-adjustment) in the ledger, so stock is
still a sum of movements. Products in the count that nobody counted keep
their stock, unless the approver chooses **Set to zero**.

Overs start unticked. An over is often a missing record (a delivery or a
return nobody entered), and the right fix is to enter that record, not to
make the stock up to the count.

## In the app

**Review** puts the counted and expected numbers side by side, and lists the
differences.

```{r}
screenshot("review.png")
```

A shortage waits under **Recount first**, with its numbers hidden until it is
recounted.

```{r}
screenshot("review-recount-first.png")
```

When the count is completed, its **Result** says what was counted, what was
short and over, and what changed. This is the demo's count at
`r showcase$location` on `r date_words(showcase$completed_at, "%B")`.

```{r}
screenshot("result.png")
```

## Why it works this way

- **Blind counts are honest counts.** Knowing the answer changes what people
  find.
- **A loss is never accepted on one count.** Shortages cost the shop money,
  and a recount is cheap.
- **Two people, not one.** One person can't count a loss away and approve it
  too. The activity log records who counted, who approved and why.
- **Short and over are kept apart.** If a shortage of
  `r rupees(-sum(showcase_short$difference * showcase_short$cost))` and an over
  of `r rupees(sum(showcase_over$difference * showcase_over$cost))` were netted
  into one figure, a real loss would hide behind something that isn't stock at
  all.
- **No percentages on small numbers.** Jholok writes "3 of 4 matched" rather
  than 75% when fewer than 10 products were counted, because a percentage of a
  handful sounds more certain than it is.

## How it works

At `r showcase$location`, the counters counted
`r nrow(showcase_lines)` products of Earrings and Rings, and
`r showcase_matched` matched: Jholok writes that as
"`r showcase_matched` of `r nrow(showcase_lines)` matched
(`r showcase_share`)". The `r tolower(showcase_short$name)`
(`r showcase_short$variant`) was one pair short. It was recounted, still one
short, and the approver accepted it: a count adjustment of
`r sub("-", minus_sign, showcase_short$difference)`. The
`r tolower(showcase_over$name)` (`r showcase_over$variant`) was one pair over. The count's note explains why: *`r tolower(showcase$notes)`*. It wasn't
the shop's stock, so the over was left unticked and the stock didn't change.

## Try it

Suppose the earrings and rings are counted at the end of the demo's story.
Unknown to anyone, a pair of diamond flower studs has gone missing. A counter
also counts one plain band twice, and finds a single chandbali earring on its
own.

```{r}
#| label: tbl-first-count
#| tbl-cap: "The products that didn't match on the first count."
first_count |>
  filter(result != "matched") |>
  transmute(
    Product = paste(name, variant, sep = ", "),
    Expected = describe_units(expected, unit),
    Counted = describe_units(counted, unit),
    Difference = paste(
      if_else(result == "short", "short", "over"),
      rupees(abs(value_at_cost)),
      "at cost"
    )
  ) |>
  table_guide()
```

`r matched_first` of `r nrow(first_count)` products matched
(`r share(matched_first, nrow(first_count))`). The studs show
short `r describe_units(abs(studs$difference), "pair")`, and the band shows
over `r describe_units(band$difference, "piece")`. The single chandbali isn't
counted as a pair: the counter counts only whole pairs, and tells the approver
about the single earring so it can be looked into.

Review asks for the studs to be recounted. The band is over, so it starts
unticked. The approver asks for it to be counted again too. On the recount,
the band matches: the over was a counting slip. The studs are still one pair
short, so the loss is real. After the recount, `r matched_recount` of
`r nrow(recount)` products match, and the approver accepts the shortage of
`r rupees(abs(studs$value_at_cost))` at cost.

**Question.** A count finds one gold coin more than expected. What should you
check before accepting it?

::: {.callout-tip collapse="true"}
## Answer

Whether a delivery or a return of coins hasn't been entered, and whether the
coin is really the shop's (a customer's, say). If a record is missing, enter
it and leave the over unticked: the ledger then matches the count without a
count adjustment.
:::

::: {.remember}
- A count compares what the ledger expects with what is really there.
- Counting is blind: counters never see the expected numbers.
- A difference is a question. A shortage is recounted before it's accepted,
  and an over usually means a missing record.
- Someone who didn't count approves, and only the differences they tick
  change the stock.
:::
