# Setting up the shop {#sec-setting-up}

```{r}
#| include: false
source(here::here("_common.R"))

demo_people <- tribble(
  ~Name, ~Role,
  "Bidisha", "Owner",
  "Sita", "Approver",
  "Ramesh", "Counter",
  "Meena", "Counter"
)

yes <- "Yes"
who_does_what <- tribble(
  ~`What`, ~Counter, ~Approver, ~Owner,
  "Count, and submit a count", yes, yes, yes,
  "Open Review and complete a count", "No",
  "Yes, if they didn't count in it",
  "Yes; if they counted in it, only when no other approver can",
  "Reopen a submitted count", "No",
  "Yes, if they didn't count in it",
  "Yes; if they counted in it, only when no other approver can",
  "Load the books into a count", "No", yes, yes,
  "Send stock out, and sign it back in", "No", yes, yes,
  "Mark a tag sold from the books", "No", "No", yes,
  "Change People, export, back up", "No", "No", yes
)
```

::: {.audience}
For the owner
:::

::: {.learn}
- What to set up before the first count, and in what order.
- How people, roles and PINs work, and why counting and approving are kept
  apart.
- How Jholok keeps the shop's stock and costs private.
:::

## The idea

Jholok works on its own from the first launch: it already knows the usual
types of jewellery and how each is counted. A few minutes of setting up makes
it fit your shop:

1. **Product types**: check that the types match how your shop groups its
   stock.
2. **Products**: add them one by one, or import a spreadsheet (@sec-catalogue).
3. **Shop details**: your shop's name and address, printed on count reports.
4. **People**: who counts and who approves, each with their own PIN.
5. **The lock**: on by default, so the app stays private.

### Owner, approvers and counters

Everyone who uses Jholok has a **role**:

- The **owner** runs the shop in Jholok. There is one owner, who sets up the
  people and proves it's them for exports and backups.
- An **approver** opens a count's Review and completes it, but only for a
  count they didn't count in. Approvers can count too.
- A **counter** counts, and submits the count for review.

This is the separation of duties from @sec-counting: one person counts,
another checks. The owner may approve a count they counted in only when
nobody else can, and the count then says so: "Counted and approved by
Bidisha (one person)".

Each counter and approver has a **PIN** of 4 to 6 digits. It shows who entered each figure,
and an approver's PIN is their signature on a completed count.

## In the app

### Product types

Open **Settings**, then **Product types**. Each type shows how many products
it has and the unit it's counted in.

```{r}
screenshot("product-types.png")
```

- Tap **+** to add a type. Give it a name, choose what it's **counted in**
  (pieces, pairs or sets), and pick a symbol.
- Tap a type to rename it or change its unit. Changing the unit doesn't
  convert existing stock: 4 pieces become 4 pairs. So choose the unit before
  adding products.
- Tap **Edit** to drag the types into the order you like.
- A type that has products can't be deleted.

### Shop details and preferences

In **Settings**, under **Preferences**:

- **Shop**: your shop's name and address. They head every count report. Add
  your GSTIN after the address if your accountant wants it on the reports.
- **Currency** and **Appearance** (light, dark, or the phone's setting).
- **Supplier lead time**: how many days a reorder takes to arrive, 21 by
  default. A product whose stock won't last that long at its recent pace of
  sales shows as running low (@sec-overview).

```{r}
screenshot("settings.png")
```

### People

Open **Settings**, then **People**. Until you set People up, Jholok asks
nobody who's counting, and counts name nobody. That's fine for trying the app
out, but set it up before your first real count.

1. Enter your name and tap **Set up People**. On a phone with Face ID, Face
   ID proves it's you; an owner PIN is optional, as a second way in. On a
   phone without Face ID, set an owner PIN.
2. Tap **Add person** for each member of staff. Enter their name, choose
   **Counter** or **Approver**, and let them choose their own PIN.

```{r}
screenshot("add-person.png")
```

```{r}
screenshot("settings-people.png", caption = "The demo shop's people.")
```

A PIN must be 4 to 6 digits. Jholok refuses easy ones: one digit repeated
(1111), a run (1234) and the most common PINs. After 5 wrong tries in a row,
the person waits a minute before trying again, and longer after each further
5 wrong tries.

To change someone's role or reset their PIN, tap their name. When someone
leaves, **Archive** them: they can no longer sign in, and the log keeps their
name. Nobody is ever deleted, so the history stays complete.

### The activity log

At the bottom of People, **Activity log** lists everything that happened:
every count entry, box, review, completion and movement, and every change to
People, each with who and when. Nothing in it can be edited or removed.

```{r}
screenshot("activity-log.png")
```

### The lock

In **Settings**, under **Security**, **Require Face ID** is on from the
start. Jholok asks for Face ID, or the phone's passcode, when it opens and
when you come back after a minute away. You can choose how soon it locks:
straight away, or after 1, 5 or 15 minutes. Whenever the app isn't in front,
a plain cover hides it from the phone's list of open apps.

Turning the lock off asks the owner to prove it's them.

## Why it works this way

- **Two people, not one.** Roles make sure the person who counts isn't the
  person who accepts a loss.
- **Once People is set up, the owner proves it's them with Face ID or the
  owner PIN, not the phone's passcode** (except to reset a forgotten owner
  PIN). On a phone shared in the shop, staff may know the
  passcode. Everything that takes data off the phone (exports, backups, box
  photos) and every change to People asks the owner each time.
- **Nobody is deleted, and the log can't be edited.** A shop's records are
  only worth something if they can't be quietly changed later.
- **The lock is on by default**, because costs and stock are private, and a
  phone left on the counter shouldn't show them.

## How it works

PINs are never stored as typed: Jholok keeps them in a scrambled form that
can't be turned back into the PIN. Each log entry carries a code made from
its own details and the code of the entry before it, so the entries form a
chain. If anyone changed or removed an entry in the phone's files, the chain
would break, and the log would say so: "The log was changed after entry 12",
instead of "intact".

If the owner forgets their PIN on a phone without Face ID, it can be reset
with the phone's passcode. The log records that, and the next count's result
says so, so a reset can't go unnoticed.

## Try it

The demo shop has four people:

```{r}
#| label: tbl-demo-people
#| tbl-cap: "Jholok Jewellers' people."
table_guide(demo_people)
```

This is who may do what:

```{r}
#| label: tbl-who-does-what
#| tbl-cap: >-
#|   Who may do what. Everything a counter does, an approver and the owner can
#|   do too.
table_guide(who_does_what)
```

**Question.** Ramesh and Sita counted the strongroom together. Who can
approve the count?

::: {.callout-tip collapse="true"}
## Answer

Not Ramesh, who is a counter. Not Sita either: she is an approver, but she
counted in this count. Bidisha, the owner, didn't count, so she can approve
it. If Bidisha had counted too, she could still approve, because no other
approver is left, and the count would say "one person".
:::

::: {.remember}
- Check the product types first: a type's unit can't convert stock later.
- Set up People before the first real count: the owner, then each counter
  and approver with their own PIN.
- Counting and approving are kept apart. The owner approves alone only when
  nobody else can.
- The lock is on from the start, and only the owner can take data off the
  phone.
:::
