# The catalogue {#sec-catalogue}

```{r}
#| include: false
source(here::here("_common.R"))

# A small spreadsheet, as a shop might keep its catalogue: two products the
# demo shop already has, each with a new price, and one new one. The band's
# stock is the same as in the app; the jhumka's is out of date.
band <- stock_on_hand(shop) |>
  filter(sku == "RIN-0001")
sample_file <- tribble(
  ~name, ~type, ~sku, ~variant, ~cost, ~price, ~on_hand,
  "Jhumka", "Earrings", "EAR-0002", "Temple Lakshmi, 22K, 14 g",
  "2,06,010", "2,45,000", "5",
  "Band", "Rings", "RIN-0001", "Plain, 22K, 4 g", "", "₹66,500",
  as.character(band$on_hand),
  "Toe ring", "Rings", "", "Silver, 4 g", "Rs. 950", "1,400", "6"
)

temple_jhumka <- stock_on_hand(shop) |>
  filter(sku == "EAR-0002")
```

::: {.audience}
For the owner
:::

::: {.learn}
- How to add a product, and what each detail is for.
- How to import a whole catalogue from a spreadsheet, safely.
- What piece tags are, and how they link to products.
:::

## The idea

The **catalogue** is the list of every product the shop sells
(@sec-stock-basics). Each product has a name, a type, a SKU, and optionally a
variant, a barcode, a cost, a list price, a photo and notes. Jholok needs the
catalogue before it can record movements or count anything.

There are two ways to build it: add products one at a time, or import them
all from a spreadsheet. Most shops start with an import, then add new designs
by hand as they arrive.

### Barcodes and piece tags

Two kinds of code can be printed on a label, and they're used differently:

- A **barcode** belongs to a product. Every piece of the same design carries
  the same barcode, just like its SKU.
- A [**piece tag**](glossary.qmd#piece-tag) belongs to one individual piece. No two pieces share a tag,
  even pieces of the same design. Billing software often prints one for each
  piece.

Because a piece tag is unique, scanning it during a count counts exactly one
piece, and scanning it twice is noticed. A barcode or SKU can't do that, so
Jholok never accepts a product's barcode or SKU as a piece tag. The [**tag
list**](glossary.qmd#tag-list) says which product each piece tag belongs to.

## In the app

### Add a product

In the **Catalogue** tab, tap **+**.

```{r}
screenshot("product-editor.png")
```

- **Photo**: choose one or take one. Without a photo, Jholok draws a picture
  of the piece from its type and name.
- **Name** and **Type**: required.
- **Variant**: what sets this design apart, such as "Temple Lakshmi, 22K,
  14 g".
- **SKU**: leave it blank and Jholok makes one from the type, such as
  EAR-0009, or type your own. No two products can share a SKU.
- **Barcode**: the product's barcode, if it has one.
- **Cost** and **Price**: what you paid for one, and what you ask. Cost puts a
  value on stock and on what a count finds short or over.
- **Opening stock**: how many the shop has now. It goes into the ledger as an
  opening stock movement. After that, stock changes only through movements:
  receipts, sales, returns, adjustments and counts.

Editing a product later never changes its stock.

### A product's page

Tap a product in the catalogue to open its page. At the top, **On hand**,
**Value** and **Cover** (@sec-stock-basics), or **In the shop**, **Off the
shelf** and **Value** when some of it is sent out (@sec-movements), then a sentence about its pace of
sales, the buttons to record a movement (@sec-movements), its stock history,
its details and every movement.

```{r}
screenshot("product-detail.png")
```

From the **⋯** menu, **Archive** a product you no longer sell: it leaves the
catalogue and the counts, but its history stays. A product with no history at
all can be deleted instead.

### Photos

Photos help people recognise products during a count. Picture matching
(@sec-boxes) compares [box photos](glossary.qmd#box-photo) only with product photos, never with
drawings. From the catalogue's
options menu, **Take photos** goes through every product without a photo,
the most valuable first. For a catalogue that looks good together, use the
same ivory [card](glossary.qmd#card) each time, in daylight, with the piece filling most of the
frame.

Where a product has no photo, Jholok draws one. During counting, a small
pencil marks a drawing, so nobody mistakes it for a photo of the piece in
their hand.

### Import a catalogue from a spreadsheet

Save the spreadsheet as a CSV file, then open **Settings** and tap **Import
catalogue from CSV**.

The file needs a **name** and a **type** column. These are optional: **sku**,
**variant**, **barcode**, **cost**, **price** and **on_hand** (the stock).
Common other headings work too, such as "Product", "Category", "Code",
"MRP" or "Qty". Amounts can be written as 1200, 1,200.50, ₹1,23,456,
Rs. 450 or 450 INR.

Nothing changes until you've seen what the import will do. The preview lists
the new products, the products it will update, and any rows it has to skip,
with the reason and the line number, so you can fix them in the spreadsheet.

### Piece tags

To link tags you already have, import a tag list: a CSV file with a **tag**
and a **sku** column, such as an export from your billing software. Open
**Settings** and tap **Import tag list from CSV**. The preview shows the new
tags and anything it can't use: tags already on another product, SKUs not in
the catalogue, and product codes that aren't piece tags.

```{r}
screenshot("settings-import-tags.png")
```

To link tags one product at a time, open the product, tap **Tags**, then
**Scan to add tags**, and hold each tag in front of the camera. A tag that
won't scan can be typed.

## Why it works this way

- **Importing never changes the stock of a product Jholok already has.** A
  spreadsheet is often out of date. If its stock differs, the preview says how
  many products differ and suggests counting them instead. While a count is
  open, it doesn't compare stock at all. Stock only changes through
  movements (@sec-ledger).
- **A blank cell never erases anything.** If the spreadsheet has no cost for
  a product that has one in Jholok, the cost stays.
- **One SKU, one product.** Jholok treats `ear-0002` and `EAR-0002` as the same,
  so a second product can't take the code by accident.
- **A piece tag is one piece.** That's what lets a scan count a piece exactly
  once.

## How it works

The import matches rows to products by SKU. A row with a SKU Jholok knows
updates that product's details. A row with a new SKU, or no SKU, makes a new
product, and its **on_hand** becomes its opening stock. A type Jholok hasn't
seen becomes a new type. Jholok guesses its unit: pairs if the name has
"earring" in it, sets if the name is or ends in "set" or "sets", and pieces
for the rest. You can change it in Settings → Product types.

## Try it

Suppose the shop imports this spreadsheet:

```{r}
#| label: tbl-sample-import
#| tbl-cap: "A small catalogue spreadsheet."
table_guide(sample_file)
```

Jholok would:

- **update the Temple Lakshmi jhumka (EAR-0002)**: its price becomes ₹2,45,000,
  but its stock stays at `r describe_units(temple_jhumka$on_hand, "pair")`,
  although the file says 5. The preview says "1 existing product has
  different stock in the file", and suggests counting it;
- **update the plain band (RIN-0001)**: its price becomes ₹66,500, and the
  blank cost leaves its cost as it was;
- **add the toe ring** as a new product, with a SKU made from its type,
  RIN-0006 (the next after RIN-0005), a cost of ₹950, a price of ₹1,400 and an opening stock of 6.

**Question.** The spreadsheet says you have 5 Temple Lakshmi jhumkas, but
Jholok says 2. What should you do?

::: {.callout-tip collapse="true"}
## Answer

Count them. If the count finds 5, the approver accepts the difference and
the ledger records a count adjustment, so the reason for the change is kept.
If it finds 2, the spreadsheet was out of date.
:::

::: {.remember}
- A product needs a name and a type. Jholok makes a SKU if you don't give
  one.
- Opening stock goes into the ledger. After that, stock changes only through
  movements.
- Importing never changes existing stock, and a blank cell never erases a
  value.
- A barcode belongs to a product; a piece tag belongs to one piece.
:::
