# Recording what comes in and goes out {#sec-movements}

```{r}
#| include: false
source(here::here("_common.R"))

# The demo with no count open, as on the screens in this chapter.
demo <- load_shop(counting = FALSE)

melted <- demo$movements |>
  filter(kind == "adjustment") |>
  left_join(select(demo$products, sku, name, variant), by = "sku")

out <- demo$out_records |>
  left_join(select(demo$products, sku, name, variant, type), by = "sku") |>
  left_join(demo$types, by = "type")
karigar <- filter(out, place == "karigar")
days_late <- as.integer(
  difftime(demo$end_at, karigar$due_back, units = "days")
)

last_30_days <- demo$movements |>
  filter(
    date > demo$end_at - 30 * 86400,
    kind %in% c("received", "sold")
  ) |>
  mutate(
    day = as.Date(date, tz = shop_time_zone),
    flow = if_else(kind == "received", "In: received", "Out: sold")
  ) |>
  summarise(units = sum(abs(quantity)), .by = c(day, flow))
```

::: {.audience}
For the owner and approvers
:::

::: {.learn}
- How to record a delivery, a sale and a return.
- How to put a wrong stock figure right.
- How to send stock out to jangad, the karigar or an order, and bring it
  back.
:::

## The idea

Between counts, the ledger stays right only if every movement is recorded:
each delivery, each sale and each return. The billing system may record
sales too; Jholok needs them as well, because its stock is the sum of its own
movements (@sec-ledger).

Three habits keep it right:

- **Record it when it happens**, or as soon as you can, with the date it
  really happened.
- **Never fix a figure by guessing.** If the stock is wrong, find out why. If
  you can't, count the product.
- **Write a note** when it will help someone later: the supplier, the bill
  number, why a piece was taken out.

### Stock that leaves the shop for a while

Not all stock is on the shelf. Jewellers often send pieces out and expect
them back:

- **Jangad**: pieces sent out on approval, with a memo, for a customer or
  another jeweller to see.
- **Karigar**: pieces with the craftsman, for repair or alteration.
- **Order**: pieces set aside for a customer's order.

These pieces are still the shop's: they count in its stock and its value. But
they aren't on the shelf, so a count mustn't expect them there. Jholok keeps
them apart: a product shows how many are **in the shop** and how many are
**off the shelf**.

```{r}
#| fig-cap: >-
#|   Where stock can be. All of it is the shop's; a count checks only the shop.
#| fig-alt: >-
#|   The shop in the middle, with arrows out to Jangad, Karigar and Order
#|   labelled Sent out, and arrows back labelled Back in, or sold there.
#| fig-height: 3.6
diagram_places()
```

## In the app

### Receive, sell and return

Open the product and tap **Receive** or **Sell**. You can also tap **+** in
the Ledger and choose the product, or **Record a sale** on the Overview.

```{r}
screenshot("record-sale.png")
```

1. Choose the kind: **Received**, **Sold** or **Returned**.
2. Set the **quantity**.
3. Add a **note** if it will help: the supplier, the bill number.
4. Check the **date and time**. If the movement happened earlier, set when it
   happened; Jholok keeps both, and the ledger shows the day it was entered
   too. A date can't be in the future.
5. **Entered by** shows who is recording it. If it isn't you, tap **Change**
   and enter your PIN.
6. Tap **Save**.

Below the quantity, Jholok shows what the stock will be. If a sale would take
it below zero, it says so, but lets you save: if the sale happened, record
it, then count the product to find what's missing.

### Put a stock figure right

If a product's stock is wrong for a reason you know (a piece damaged and
sent for melting, say), open it, tap **Adjust**, and enter the **correct
stock**. A **reason** is required. Jholok records the difference as an
adjustment, with your reason, in the ledger.

In the demo, a `r tolower(melted$name)` (`r melted$variant`) was taken out
on `r date_words(melted$date, "%B")`: "`r melted$note`".

Adjust is turned off for a product while a count covers it, until the count
is reviewed. Use the count to put its stock right instead.

### Send stock out, and bring it back

Open the product and tap **Send out**. An approver or the owner signs it
with their PIN.

```{r}
screenshot("send-out.png")
```

- **To**: Jangad, Karigar or Order.
- **How many**, or choose the pieces by their tags.
- **Party**: who has it, and **Memo**: the memo number.
- **Due back**: when it should come back. Jholok suggests 7 days for jangad
  and 30 for the karigar, and none for an order; change it if you agreed
  otherwise.

The product's page then shows an **Off the shelf** section, with each memo,
who has it and when it's due. When the pieces come back, tap the memo and
choose **Back in**. If the customer keeps them, choose **Sold from here**.
Both are signed by an approver or the owner.

A memo more than a week overdue appears under **Needs attention** on the
Overview (@sec-overview).

## Why it works this way

- **A movement can't be edited or deleted.** A mistake is put right with
  another movement, usually an Adjust with its reason, so the ledger keeps
  the whole story (@sec-ledger).
- **The date is when it happened; Jholok also keeps when it was entered.**
  A sale entered late still lands on the right day, and anyone can see it was
  entered late.
- **A sale below zero is allowed.** Refusing it would leave a real sale
  unrecorded. Jholok records the truth and points to the missing piece.
- **Adjust waits while a product is being counted.** Changing the ledger in
  the middle of a count would change what the count is checked against.
- **Sending out needs a second person.** Stock leaving the shop is when
  pieces go missing, so an approver or the owner signs for it, and every
  memo is logged.

## How it works

Each movement stores its date, its kind, its quantity and note, who entered
it and when it was entered. An Adjust stores the difference between the
stock before and the correct stock you entered.

Sending out is two movements that add up to nothing: the stock leaves the
shop and arrives at the place. So the stock the shop owns, and its value,
don't change, while the stock in the shop goes down. Back in moves it the
other way. Sold from here records a sale from that place.

If you record a movement dated before a count that has already been
completed, Jholok checks whether that count already accounted for it, so it
isn't counted twice. Chapter 12 explains this.

## Try it

This is what came in and went out of Jholok Jewellers in the last 30 days,
as the Ledger tab sums it up:

```{r}
#| fig-cap: >-
#|   Pieces, pairs and sets received and sold each day in the last 30 days.
#| fig-alt: >-
#|   Two rows of bars by day: received above, sold below, on the same scale.
#| fig-height: 3.2
ggplot(last_30_days, aes(day, units)) +
  geom_col(aes(fill = flow), width = 0.7) +
  facet_wrap(vars(flow), ncol = 1) +
  scale_fill_manual(
    values = c(
      "In: received" = guide_colours$jade,
      "Out: sold" = guide_colours$gold_deep
    ),
    guide = "none"
  ) +
  scale_x_date(date_labels = "%d %b", breaks = scales::breaks_pretty(5)) +
  labs(x = NULL, y = NULL) +
  theme_guide() +
  theme(strip.text = element_text(hjust = 0, face = "bold"))
```

And this is the stock the demo has out of the shop:

```{r}
#| label: tbl-out-of-shop
#| tbl-cap: "Stock off the shelf in the demo shop."
out |>
  transmute(
    Product = paste(name, variant, sep = ", "),
    Where = paste0(toupper(substr(place, 1, 1)), substring(place, 2)),
    How_many = describe_units(units, unit),
    Party = party,
    Memo = memo,
    `Due back` = if_else(is.na(due_back), "–", date_words(due_back))
  ) |>
  rename(`How many` = How_many) |>
  table_guide()
```

The chain with the karigar was due back `r days_late` days ago, so it is
listed under Needs attention.

**Question.** A delivery of 3 pairs arrived on Monday, but you enter it on
Wednesday. What date do you give it?

::: {.callout-tip collapse="true"}
## Answer

Monday, when it arrived. Jholok keeps Wednesday too, as the day it was
entered, and the ledger shows both.
:::

::: {.remember}
- Record every delivery, sale and return, dated when it happened.
- A movement is never edited. Put a wrong figure right with Adjust, and give
  the reason.
- Adjust waits while the product is being counted.
- Stock on jangad, with the karigar or kept for an order is still the shop's,
  but a count doesn't expect it on the shelf. An approver or the owner signs
  it out and back.
:::
