# Reading the Overview {#sec-overview}

```{r}
#| include: false
source(here::here("_common.R"))

# The demo with no count open, as on the Overview screens here.
demo <- load_shop(counting = FALSE)

# The 6-month view: 26 weeks, the current one included.
window_start <- week_of(demo$end_at) - 25 * 7
before_window <- as.POSIXct(
  paste(window_start - 1, "23:59"),
  tz = shop_time_zone
)

type_now <- value_by_type(demo)
type_then <- value_by_type(demo, before_window)
type_change <- type_now |>
  left_join(type_then, by = "type", suffix = c("_now", "_then")) |>
  mutate(change = value_now - value_then)
total_now <- sum(type_change$value_now)
total_change <- sum(type_change$change)
biggest <- slice_max(type_change, abs(change), n = 1)

sales <- weekly_sales(demo) |>
  filter(week >= window_start)
complete_weeks <- filter(sales, week < week_of(demo$end_at))
last_week <- slice_tail(complete_weeks, n = 1)
busiest <- slice_max(complete_weeks, value, n = 1)
busiest_name <- if (is.na(busiest$festival)) {
  paste("the week of", date_words(busiest$week))
} else {
  paste0(busiest$festival, " week, ", date_words(busiest$week))
}

attention <- attention_lists(demo)
lists <- count(attention, list)
```

::: {.audience}
For the owner
:::

::: {.learn}
- What each part of the Overview tells you.
- What makes a product "need attention", and what to do about it.
- How to read the figures honestly: what they include, and what they leave
  out.
:::

## The idea

The **Overview** is the first tab: a summary of the whole shop on one
screen. It answers the owner's daily questions. What is my stock worth, and
is that going up or down? Which types hold the money? What's selling, and
what isn't? What needs doing? How did the last count go?

Every figure comes from the ledger (@sec-ledger), worked out afresh when you
open the app. Nothing on the Overview can be typed in.

At the top, **3M**, **6M** and **1Y** choose how far back the charts look:
13, 26 or 52 weeks, this week included. Needs attention always looks at the
last 90 days.

## In the app

### Stock value

```{r}
screenshot("overview.png")
```

The large figure is what the stock is worth at cost,
`r rupees_short(total_now)` "at today's cost". The sentence under it says how
that changed over the period, and which type changed most: in the demo,
"Down `r rupees_short(abs(total_change))`
(`r percent_words(total_change / (total_now - total_change))`) since
`r date_words(window_start)`, mostly
`r biggest$type` (`r rupees_short(biggest$change)`)". The line shows the value
week by week.

Below it: the value at list price, what's in stock in each unit, and how many
products there are. The small print says what the figure leaves out, such as
products with no cost, and that past values use today's costs, so the line
shows changes in stock, not in the price of gold.

### By type, Needs attention, Sales, Best sellers and Last count

```{r}
screenshot("overview-lower.png")
```

- **By type** shows how the value is shared between types, and how each has
  moved. Tap it for a chart of each type on the same scale.
- **Needs attention** lists products that need something done (below).
- **Sales** shows the value sold each week at list price, after returns.
- **Best sellers** lists the five products that sold the most by value.
- **Last count** says how the last count went, and which types have gone
  longest without one.

### Needs attention

```{r}
screenshot("overview-attention.png")
```

Each list has a rule, and something to do about it:

| List | Means | What to do |
|---|---|---|
| Below zero | The ledger shows less than none: a movement is missing | Record the missing delivery, or count it |
| Out of stock, still selling | None left, but it sold in the last 90 days | Reorder |
| Running low | At its recent pace, the stock won't last the supplier lead time (21 days unless you change it) | Reorder soon |
| Not selling | In stock, with no sale in 90 days | Consider a discount, a display change, or melting |
| No cost, No price | In stock with no cost or no list price, so values leave it out | Add the missing amount |
| Overdue memos | Stock on jangad or with the karigar more than a week past its due date | Chase it up, or record it back in |

A product only just added (in the last 30 days) isn't called low or not
selling: it hasn't had time to show its pace.

### Sales

```{r}
screenshot("overview-sales.png")
```

Each bar is a week, Sunday to Saturday. The last complete week is gold. The
busiest week is labelled with its festival, when there was one.

## Why it works this way

- **Values say their basis and date** ("at today's cost, 8 Oct"; "at today's
  list price"), because a figure without them can mislead.
- **Sales are valued at today's list price, not the price charged.** Jholok
  doesn't record discounts, so this shows how much stock sold, not the
  takings. Your billing system has the takings.
- **Nothing is quietly left out.** Products with no cost or price, and types
  being counted, are named in the small print.
- **While a count is open, the stock of the types being counted is hidden**,
  with a note at the top: "Stock for Necklaces is hidden while the count at
  Showcase 2 is open." Sales, Best sellers and Last count stay.

## How it works

The value line takes each product's stock on each day, from the ledger, and
values it at today's cost. Archived products are left out. A product's
**pace** is the number it sold in the last 90 days, after returns, divided by
the days it was in stock (at least 14), and its **cover** is how many days
its stock lasts at that pace. Running low means cover shorter than the
supplier lead time, which you can change in Settings → Preferences.

The sales figures add up every unit sold, whatever its kind, and call them
**items**: "9 items" can be pairs, sets and pieces together.

## Try it

These are the demo shop's sales over the last six months. Everything here is
worked out from the ledger, the same way as the app.

```{r}
#| fig-cap: >-
#|   Sales by week at today's list price, after returns. The festival weeks are
#|   labelled.
#| fig-alt: >-
#|   Weekly bars over six months, highest in the Akshaya Tritiya week in
#|   early May.
#| fig-height: 3.4
peak <- slice_max(complete_weeks, value, n = 1)
ggplot(sales, aes(week, value)) +
  geom_col(
    aes(fill = week == last_week$week),
    width = 5
  ) +
  geom_text(
    aes(label = festival),
    data = peak,
    vjust = -0.6,
    size = 3.4,
    colour = guide_colours$text_secondary
  ) +
  scale_fill_manual(
    values = c(
      `TRUE` = guide_colours$gold_mid,
      `FALSE` = guide_colours$neutral
    ),
    guide = "none"
  ) +
  scale_y_rupees(expand = expansion(mult = c(0, 0.15))) +
  scale_x_date(date_labels = "%b") +
  labs(x = NULL, y = NULL) +
  theme_guide()
```

Last week the shop sold `r rupees_short(last_week$value)`,
`r last_week$items` items. The busiest week was `r busiest_name`, with
`r rupees_short(busiest$value)`.

And these are the products that need attention:

```{r}
#| label: tbl-attention
#| tbl-cap: >-
#|   What needs attention in the demo shop. One product can be on two lists.
attention |>
  transmute(
    List = as.character(list),
    Product = paste(name, variant, sep = ", "),
    SKU = sku
  ) |>
  table_guide()
```

That's `r n_distinct(attention$sku)` products, as the Overview says.

**Question.** A pendant shows under **Not selling**. Does that mean it's
been stolen?

::: {.callout-tip collapse="true"}
## Answer

No. It means it's still in stock but hasn't sold in 90 days. It may be the
wrong design for this season, or not on display. Only a count can tell you
whether stock is missing.
:::

::: {.remember}
- The Overview is worked out from the ledger every time; nothing on it is
  typed in.
- Every value says what it's based on: today's cost or today's list price.
- Needs attention says what to do: reorder, chase a memo, add a cost, or
  count.
- While a count is open, the stock being counted is hidden here too.
:::
