# A count from start to finish {#sec-a-count}

```{r}
#| include: false
source(here::here("_common.R"))

# Showcase 2, the demo's open count of the necklaces.
open_count <- shop$counts |>
  filter(is.na(completed_at))
in_scope <- shop$products |>
  filter(
    type %in% unlist(open_count$scope),
    created_at <= open_count$started_at,
    is.na(archived_at)
  )
counted_so_far <- shop$count_lines |>
  filter(count_id == open_count$count_id) |>
  left_join(select(shop$products, sku, name, variant), by = "sku")
```

::: {.audience}
For counters
:::

::: {.learn}
- How to start a count, and choose what it covers.
- How to count each product quickly, with one hand, and put mistakes right.
- How to submit the count when you're done.
:::

## The idea

@sec-counting explained why counts matter. This chapter is the counter's side:
the steps on the phone, from starting a count to handing it over for review.

Three things are worth knowing before you start:

- **A count can be small.** It can cover the whole shop, a few types, or the
  pieces in one showcase. Smaller, more frequent counts find problems sooner.
- **You count blind.** Nothing on the counting screen shows how many there
  should be. While the count is open, the rest of the app hides the stock of
  the types being counted too.
- **Every tap is saved at once.** If the phone rings or the app closes, the
  count is still there when you come back.

## In the app

### Start a count

In the **Count** tab, tap **New count** (the + at the top).

```{r}
screenshot("new-count.png")
```

- **Location**: where you're counting, such as "Front case". The places used
  in recent counts are offered below the field, so you can tap one.
- **What to count**: leave **All types** on to count everything, or turn it
  off and tick the types you want. Each type says when it was last counted.
- **Notes**: anything the approver should know. It's optional.

The line under the types says how many products the count covers, and
reminds you that "Expected stock stays hidden until you review, so it can't
sway the count." Tap **Start**.

The Count tab also suggests what to count next: under **Last counted**, a
button such as "Count Sets" starts a count of the type that has gone longest
without one.

```{r}
screenshot("counts.png")
```

### Say who's counting

If the owner has set up People (@sec-setting-up), the app asks **Who's
counting?** Choose your name if it isn't already chosen, and enter your PIN.

```{r}
screenshot("who-is-counting.png")
```

You stay signed in while the count is open, even if you leave the app for up
to 15 minutes. After that, or if you go back to the list of counts, the app
asks again, with your name already chosen. Several people can count in one
count: each signs in with their own name, and the count lists everyone who
counted.

### Count each product

The counting screen lists the products in the count, grouped by type. The
card at the top shows how far you've got, such as "4 of 9 products counted".

```{r}
screenshot("counting.png")
```

Each product has three controls:

- **+** adds one. Tap it once for each piece, pair or set as you pick it up.
  It sits on the right, under your thumb.
- **−** takes one away, if you tapped once too often.
- **The number** shows your count. A dash (–) means the product hasn't been
  counted yet. Tap the number to type a count instead.

When you tap the number, choose **Add** to add to what you've counted so far
(useful when you count tray by tray), or **Set total** to type the whole
count. The sheet shows the count so far and the new total before you tap
**Done**.

```{r}
screenshot("quantity-entry.png")
```

Two quick actions are hidden behind a swipe on a product:

- **Swipe right for None here**, when you've looked and there are none. This
  counts 0, which is different from not counting it at all.
- **Swipe left for Clear**, to take your count off and start again. It's
  there once you've counted the product yourself.

A swipe only shows the button: you still have to tap it. That way a stray
swipe while your other hand holds a tray never changes a count.

### Put a mistake right

After any change of more than one, and after every Clear, a bar at the bottom
offers **Undo** for a few seconds. It says what changed, such as "Chain set to
7 pieces, was 9 pieces." Tap Undo to put the count back exactly as it was.

```{r}
screenshot("undo-banner.png")
```

### Find a product quickly

Type in **Find a product** to search by name, SKU or barcode. The filter
button at the top shows only the products **Not counted**, or only those
**Counted**, which helps you check nothing was missed.

To find a product from its label, tap the barcode button, **Find by code**.
Scan or type a barcode, a SKU or a piece tag, and the list jumps to that
product's row. Finding a product never counts it: you still tap +.

```{r}
screenshot("find-by-code.png")
```

The buttons at the bottom, **Scan tags** and **Box photo**, count a whole tray
or box at once. Chapter 9 explains them.

### Submit the count

When you've counted everything, tap **Submit**. The app checks that you mean
it.

```{r}
screenshot("submit-dialog.png")
```

After you submit, counting stops and the count waits for an approver. Nobody
can change a figure unless an approver reopens the count, with a reason.

```{r}
screenshot("waiting-for-approval.png")
```

## Why it works this way

- **One hand is enough.** Counting happens with a tray in the other hand. The
  + is large and stays at the right edge, under the thumb, and the phone gives
  a small tap of feedback with every change.
- **No swipe changes a count by itself**, for the same reason.
- **Nothing shows stock while you count.** Not the counting screen, and not
  the rest of the app either: the Overview and the Catalogue show "Being
  counted" for the types in an open count.

```{r}
screenshot(
  "overview-blind.png",
  caption = "The Overview while the necklaces are being counted."
)
```

- **None here is not the same as not counted.** "None here" says you looked
  and found nothing. A product left with a dash wasn't counted, and keeps its
  stock unless the approver decides otherwise.
- **Submitting stops counting** so that the approver reviews a count that
  can't change under them.

## How it works

Your count is saved with every tap, together with who entered it and when.
The activity log records each figure, so if several people count, it's clear
who counted what. Pieces counted from box photos or tag scans add to the same
number as your taps; chapter 9 explains how the two are kept apart.

When you submit, the count waits under **In progress** in the Count tab,
marked "Waiting for approval". An approver then reviews it, as @sec-counting
explained.

## Try it

The demo's open count is at `r open_count$location`. It covers the
`r tolower(unlist(open_count$scope))`: `r nrow(in_scope)` products. So far,
`r nrow(counted_so_far)` have been counted:

```{r}
#| label: tbl-counted-so-far
#| tbl-cap: >-
#|   The necklaces counted so far at Showcase 2. Like the counter, this table
#|   shows no expected stock.
counted_so_far |>
  transmute(
    Product = paste(name, variant, sep = ", "),
    Counted = describe_units(counted, "piece"),
    `Counted at` = format(counted_at, "%H:%M", tz = shop_time_zone)
  ) |>
  table_guide()
```

The other `r nrow(in_scope) - nrow(counted_so_far)` still show a dash.

**Question.** You've counted 3 rope chains on the showcase, then find 2 more
in the drawer below. What's the quickest way to record them, and what will the
app show?

::: {.callout-tip collapse="true"}
## Answer

Tap +, twice. Or tap the number, choose **Add**, type 2 and tap **Done**: the
new total is 5, and because the count changed by more than one, the app
offers Undo for a few seconds.
:::

::: {.remember}
- Start a count from the Count tab, and choose where and what it covers.
- Tap + for each piece. Tap the number to type a count. Swipe for None here or
  Clear, then tap.
- Undo is offered after a change of more than one, and after every Clear.
- Submit when you're done; after that, only an approver can reopen the count.
:::
