# Counting boxes: tags and photos {#sec-boxes}

```{r}
#| include: false
source(here::here("_common.R"))

boxes <- sample_boxes()
tagged <- filter(boxes, box == 1)
outcomes <- box_tag_outcomes(1)
by_outcome <- count(outcomes, outcome)
outcome_n <- \(name) sum(by_outcome$n[by_outcome$outcome == name])
to_identify <- outcome_n("no tag read") + outcome_n("tag not in the tag list")
unknown_tag <- outcomes |>
  filter(outcome == "tag not in the tag list") |>
  pull(tag)
by_tag <- outcome_n("identified by its tag")
tag_list <- demo_tag_list()
```

::: {.audience}
For counters
:::

::: {.learn}
- The two ways to count a whole box at once: scanning its tags, or taking a
  photo.
- How to check a box before adding it to the count.
- What the app never does on its own, and why.
:::

## The idea

Counting a box of 15 pairs one tap at a time works, but it's slow. Jholok can
count a whole box at once, in one of two ways:

- **Scan tags.** If each piece carries its own piece tag (@sec-catalogue),
  move the camera over the box. Each tag read counts its piece, once. This is
  the most reliable way, which is why **Scan tags** is the gold button.
- **Box photo.** Photograph the box. Jholok finds the box, straightens it,
  finds each card, reads any tags it can see, and suggests a product for the
  rest from their pictures.

A box holds many designs of one type, each pair or set on its own card or in
its own slot. So the app counts **cards**, not loose pieces: one card is one
pair, or one set.

Either way, the result is a [**proposal**](glossary.qmd#proposal). The review says "15 pairs
proposed": nothing reaches the count until you've checked it and tapped
**Add**.

::: {.callout-important}
## How well box photos work hasn't been measured yet

The app's photo counting was built and tuned on drawn example boxes, like the
ones in this guide, not on photos of real boxes. How often it gets a real
shop's boxes right isn't known yet. That's why every box is checked before
it's added, and why the app never adds a box by itself.
:::

```{r}
#| fig-cap: >-
#|   The app counts cards. Where a card may hold two, or cards touch, it
#|   asks you; it never splits a piece on its own.
#| fig-alt: >-
#|   Three drawings: one card with a pair, counted 1; two overlapping cards
#|   marked Make it 2?; two touching cards marked with a warning.
#| fig-height: 2.6
diagram_cards_not_pieces()
```

## In the app

On the counting screen, the two buttons at the bottom open the camera:
**Scan tags** and **Box photo**. The camera button beside each type's name
photographs a box of that type.

```{r}
screenshot("counting-strongroom.png")
```

### Scan tags

1. Tap **Scan tags**, and move the camera slowly over the tags. It reads
   several at once.
2. Each tag read joins the list with what happened to it. The phone gives a
   different buzz when a tag is read again, so you know to move on.
3. For a tag the camera can't read, tap **Type a tag** and type its code. It
   counts once, like a scanned tag.
4. Tap **Done** to see the summary.

| You see | It means |
|---|---|
| +1 | Counted |
| Already counted in Box 2 | This piece was counted earlier in the count, so it adds nothing |
| Not in your tag list | The tag isn't linked to a product yet |
| Product code | A barcode or SKU, which is on every piece of a design, so it can't count one |
| Not in this count | A product this count doesn't cover |
| Archived | A product no longer in the catalogue |
| Customer's piece, not stock | A customer's own piece, which is never the shop's stock |

The summary says what the scan found. A tag the list doesn't know must be
linked to its product (**Link…**) or skipped before you can add the scan.
Linked tags join your tag list when you add it.

```{r}
screenshot("box-tags-1.png")
```

Link or skip the unknown tag, then tap **Add … pairs to count**: in this
scan, `r by_tag + 1` if you link it, `r by_tag` if you skip it. The scan is kept as "Tag scan 1", and each
counted tag adds one to its product.

### Take a box photo

1. Tap **Box photo**. If the count covers several types, choose the box's
   type at the top.
2. Hold the phone flat above the box, with the whole box in the frame. If
   it's tilted, the app says "Hold the phone flat". Tags it can read are
   ticked as you frame.
3. Tap the shutter. The app checks the photo, and if it's blurred, too dark,
   too bright or tilted, offers **Retake** or **Use anyway**.
4. It then finds the pieces, reads the tags and matches pictures, and opens
   the box's review.

```{r}
screenshot(
  "box-capture.png",
  caption = paste(
    "Taken on a test phone with no camera. Your phone shows the camera view",
    "here, and has no Sample boxes button."
  )
)
```

You can also **Choose photo** to use one you took earlier.

### Check the box

The review shows the photo with every piece outlined and numbered. A small
mark on each outline says how its product was found:

- a **tag**: read from its tag;
- **sparkles**: suggested by its picture, waiting for you to check it;
- a **question mark**: still needs a product;
- no mark: confirmed.

A **dashed** outline is a piece left out, and a red **!** is a piece that may
be two, or was found only by its tag.

```{r}
screenshot("box-review-1.png", caption = drawn_box)
```

Tap any piece to open it. The sheet shows its tag, its product, and the
products it may be, with a search at the top. From there:

- **choose the product**, or **accept** the suggested one;
- **Make it 2**, on a piece marked with a red **!** as maybe two. If one
  outline holds two pieces and isn't marked, use **Add piece**;
- [**Not a piece**](glossary.qmd#not-a-piece), when the outline isn't jewellery (an empty card, a
  label). You can undo it until your next change;
- [**Not in the catalogue**](glossary.qmd#not-in-the-catalogue), for a design the shop has no product for;
- **Identify later**, on a picture match you can't check now: it clears the
  match, and the piece waits in the [To identify](glossary.qmd#to-identify) queue
  (@sec-identify-later). A piece with no product goes there anyway.

```{r}
screenshot("piece-sheet.png", caption = drawn_box)
```

If the app missed a piece, tap **Add piece**, then tap where it is. For a tray
with one piece per slot, such as rings, turn on **Count by tray grid**, set
the rows and columns, and drag the corners to fit the tray.

```{r}
screenshot("add-piece.png", caption = drawn_box)
```

When everything looks right, tap **Add 15 pairs to count**. Any piece still
without a product goes to the To identify queue. A piece whose picture match
you haven't checked keeps the box from being added: accept it, change it, or
choose Identify later first.

### If the box was counted already

If a box photo shares half or more of its tags with a box already added, or
an untagged box looks just like one already added, the app asks first: "Is
this Box 2 again?" If it is, tap **Cancel**, or its pieces would be counted
twice.

### After adding

The count lists the box under **Boxes**, and its pieces add to each product's
count, together with anything counted by hand. Tap a box to open its review
again and update it, or swipe it to **Remove** it: its pieces come off the
count, and its photo is deleted.

```{r}
screenshot("counting-after-box.png")
```

## Why it works this way

- **Tags first.** A piece tag names one piece exactly, so scanning can't
  mistake one design for another.
- **A tag counts once per count.** Whether it's scanned twice, or appears in
  a photo and a scan, the piece is counted once.
- **Cards, not loose pieces, and never a guessed split.** Earrings lie close
  together and sets are large; guessing would be wrong too often. Where one
  outline may hold two, the app asks; where pieces touch, it warns.
- **A proposal, never automatic.** The app can't yet show it counts real boxes
  well, so a person checks every box. Settings → Photo counting keeps score.
- **Not a piece is kept, left out.** What the app got wrong is exactly what a
  better version needs to learn from, so the outline stays with the box,
  marked, and is never counted.
- **Picture matching never looks at stock.** It compares the piece's picture
  with product photos of the box's type only.

## How it works

Every added box keeps its photo, the straightened box, and every piece you
confirmed. These are kept so that, in future, the app can learn from your
real boxes. The owner can export them from Settings → Export → Box photos and
labels, which asks the owner to confirm each time.

Settings → Photo counting compares the app's count of each box with the
count you added after checking it. Those numbers lean towards the app,
because you saw its proposal first, so they aren't an independent measure.

```{r}
screenshot("settings-photo-counting.png")
```

## Try it

This is the demo's tagged box of earrings, with each piece numbered as in
its review.

```{r}
#| fig-cap: !expr box_caption("The tagged box of earrings.")
#| fig-alt: >-
#|   A maroon velvet box of 15 pairs of earrings on white cards, each outlined
#|   and numbered, with one empty card.
#| fig-height: 6
plot_box_photo(1)
```

Its `r tagged$pieces` pieces carry tags, but a photo of it doesn't read them
all:

```{r}
#| label: tbl-box-one-tags
#| tbl-cap: "What happens to each piece's tag in the tagged box."
by_outcome |>
  transmute(
    `What happened` = paste0(
      toupper(substr(outcome, 1, 1)),
      substring(outcome, 2)
    ),
    Pieces = n
  ) |>
  arrange(desc(Pieces)) |>
  table_guide()
```

`r outcome_n("identified by its tag")` pairs are identified by their tags.
The tag on one pair, `r unknown_tag`, isn't among the `r nrow(tag_list)` tags
in the demo's tag list: that pair arrived after the list was exported, so the
review asks to link it. Two tags can't be read. That leaves `r to_identify`
pairs to identify, as the review says: "`r to_identify` to identify later,
from the count."

The empty card in the bottom corner holds no jewellery. If the app had
outlined it, you would tap it and choose **Not a piece**.

**Question.** You scan a box and the list shows "Already counted in Box 2"
for every tag. What happened?

::: {.callout-tip collapse="true"}
## Answer

This box was already counted, as Box 2, earlier in the count. Its pieces add
nothing, so nothing is counted twice. Don't add the scan.
:::

::: {.remember}
- Scan tags when pieces have tags; take a box photo when they don't.
- Every box is a proposal: check it, then Add. Nothing is added by itself.
- The app counts cards, asks when one may be two, and never splits a piece.
- A tag counts once per count. If the app asks whether a box was counted
  already, believe it until you've checked.
- How well box photos work on real boxes hasn't been measured yet.
:::
