# Pieces to identify later {#sec-identify-later}

```{r}
#| include: false
source(here::here("_common.R"))

untagged <- sample_boxes() |>
  filter(box == 2)
box_two <- sample_box_counts(2, shop)
most_common <- slice_head(box_two, n = 1)
```

::: {.audience}
For counters
:::

::: {.learn}
- Why a box photo can be added before every piece in it is identified.
- How to identify pieces from the To identify queue.
- Why the count can't be completed until every piece is identified, and why
  only counters identify.
:::

## The idea

When a box's pieces carry tags, Jholok knows each piece's product from its
tag (@sec-boxes). In a box with no tags, a photo can still find the cards and
count them, but it doesn't know which design is on each card. Someone has to
say.

Doing that piece by piece in the box review is slow: often slower than
counting by hand. So Jholok lets you **add the box first** and **identify its
pieces later**, from one list, the **To identify** queue.

Until a piece is identified, it counts in the box but in no product. The box
adds "15 pairs", but no product's count goes up for that piece yet. And the
count can't be completed until every piece is identified, or marked as not in
the catalogue. That way, no product is ever called short just because nobody
said which pieces were its.

## In the app

### Add the box

In the review of an untagged box, every piece without a product is marked
for identifying later. The gold button still adds the box, and the note under
it says how many pieces are left: here, all `r untagged$pieces`.

```{r}
screenshot("box-review-2.png", caption = drawn_box)
```

Tap **Add `r untagged$pieces` pairs to count**. The count now shows a **To
identify** row at the top, and the type's heading says how many of its pieces
are waiting.

```{r}
screenshot("counting-to-identify.png")
```

### Identify each piece

Tap the **To identify** row. The queue shows one piece at a time.

```{r}
screenshot("identify-queue.png", caption = drawn_box)
```

- At the top: the piece's picture, how many are left ("1 of 15" means 15
  still to identify), and which
  box and piece it is. **See the box** shows the whole photo with this piece
  outlined, so you can find it in the tray.
- Below: a search field, then **picture matches** (the products whose photos
  the piece looks most like), then every product of the box's type, as
  pictures. Picture matching compares with product photos only, so products
  need photos for it to help.
- **Tap the product** the piece is. The queue moves to the next piece.
- **Skip** leaves a piece for later. When the rest are done, the queue says
  how many you skipped; tap **Start again** to go back to them.
- **Not in the catalogue**, at the bottom, is for a design the shop has no
  product for. The piece is left out of the count, and the count's result and
  report list it by box and piece number, so someone can add the product and
  count it later.

After you pick a product, if other pieces in the same box look the same, the
queue offers them too, ticked: "These 3 in Box 1 look the same. Untick any
that aren't." Untick any that differ, then tap **Identify 4 as …**, or **Just
piece 1** to identify only the first. Nothing is identified by its look
alone: you always confirm.

### Submit with pieces left

You can submit a count with pieces still to identify. The app offers
**Identify now** first, and **Submit anyway** second, and says the count can't
be completed until they're done. The approver's Review then says how many
pieces aren't identified, and **Complete** stays off until they are. The
approver reopens the count, with a reason, and a counter identifies them.

## Why it works this way

- **It's quicker.** One list for every box, one tap per piece, and lookalikes
  in one go, instead of opening each piece in each box.
- **Unidentified isn't missing.** A piece nobody has identified still counts
  in its box. Completing the count with it would call its product short, so
  completing waits.
- **Only counters identify, while counting.** The approver can see what's
  short. If the approver could say which product a piece was, a loss could be
  covered up by "finding" the missing product in a box. So the approver
  reopens the count, and a counter, who can't see stock, identifies the
  pieces.
- **The queue is blind.** It shows pieces and products, never stock.
- **A design not in the catalogue doesn't block the count forever.** It's left
  out, and listed, so it gets dealt with.

## How it works

Each identification is saved at once, with who made it and when, in the
activity log. The piece's product is set, and that product's count goes up.
Identified pieces also help picture matching: until the app is closed, a
piece like it in another photo is suggested that product.

Lookalikes are found by comparing the pieces' pictures with each other,
never with the stock. This needs a real phone: on the test phones used for
this guide's screenshots, the queue never offers lookalikes.

Like everything about box photos, picture matching was tuned on drawn
example boxes. How well it does on a real shop's boxes hasn't been measured
yet: always check the product before you tap it.

## Try it

This is the demo's untagged box. Its pieces have no tags, so every piece is
waiting to be identified.

```{r}
#| fig-cap: !expr box_caption("The untagged box, numbered as in the queue.")
#| fig-alt: >-
#|   A navy velvet box of 15 pairs of earrings on cream cards, each outlined
#|   and numbered.
#| fig-height: 6
plot_box_photo(2)
```

If every piece is identified correctly, the box adds these to the count:

```{r}
#| label: tbl-box-two
#| tbl-cap: "What the untagged box adds, once its pieces are identified."
box_two |>
  transmute(
    Product = paste(name, variant, sep = ", "),
    Pairs = pieces
  ) |>
  table_guide()
```

Several pieces share a design, such as the `r tolower(most_common$name)`
(`r most_common$variant`), with `r most_common$pieces` pairs in this box.
On a phone, identifying one of them may offer the others as lookalikes, if
their pictures are close enough.

**Question.** A piece in the box is a design the shop bought last week but
hasn't added to the catalogue yet. What do you do?

::: {.callout-tip collapse="true"}
## Answer

Tap **Not in the catalogue**. The piece is left out of this count, and the
count's result lists it by box and piece number. Then the owner adds the
product to the catalogue, and the piece can be counted in the next count.
:::

::: {.remember}
- An untagged box can be added before its pieces are identified.
- Identify them from the To identify queue: one tap per piece, and
  lookalikes together.
- A design with no product is marked Not in the catalogue, and listed.
- The count can't be completed until every piece is done, and only counters,
  while counting, identify.
:::
