# Stock off the shelf and the cut-off {#sec-cut-off}

```{r}
#| include: false
source(here::here("_common.R"))

# The demo with the karigar's chain sent out inside the open count, as on
# the Review screenshot in this chapter.
off_shelf <- load_shop(off_shelf_in_count = TRUE)
open_count <- filter(off_shelf$counts, is.na(completed_at))
out <- off_shelf$out_records |>
  left_join(select(off_shelf$products, sku, name, variant), by = "sku")
chain_line <- off_shelf$count_lines |>
  filter(count_id == open_count$count_id, sku == out$sku)
chain_in_shop <- stock_in_shop(off_shelf, chain_line$counted_at) |>
  filter(sku == out$sku) |>
  pull(in_shop)
chain_owned <- stock_on_hand(off_shelf, chain_line$counted_at) |>
  filter(sku == out$sku) |>
  pull(on_hand)
```

::: {.audience}
For the owner and approvers
:::

::: {.learn}
- How a count treats stock that is the shop's but not on the shelf.
- What the cut-off is, and why sales during a count don't look like losses.
- What happens when a movement is entered after its count was completed.
:::

## The idea

A count compares the shelf with the ledger at one moment. Two things make
that moment hard to pin down in a real shop: some stock is away, and the shop
stays open while it's counted.

### Stock that is away

Pieces on jangad, with the karigar, or kept for an order are the shop's stock,
but they aren't on the shelf (@sec-movements). A count expects only what's
**in the shop**: owned stock minus what's away. The rest is listed apart, so
nobody mistakes it for a loss.

Customers' own pieces, in for repair, are never the shop's stock, wherever
they sit.

### The cut-off

A count takes time. If a pair of jhumkas is sold at 9:30, and the jhumkas
were counted between 9:05 and 9:40, which stock should the count be compared with:
before the sale, or after?

Jholok works out a [**cut-off**](glossary.qmd#cut-off) for every product in a count. It reads the
activity log to find when that product was counted (its [**counting window**](glossary.qmd#counting-window)),
and compares the count with the stock in the shop at the end of that window.
If stock moved during the window, it checks every stock figure along the
way: if the count matches any of them, the count is fine. If it matches none,
the product needs a recount.

```{r}
#| fig-cap: >-
#|   A sale during the counting window. A count that matches the stock before
#|   or after the sale is fine; one that matches neither is recounted.
#| fig-alt: >-
#|   A stock line at 5 that drops to 4 at 9:30, inside a shaded counting
#|   window. Counts of 5 and 4 are marked fine, a count of 3 is marked recount.
#| fig-height: 3.2
diagram_cut_off()
```

So the shop never has to close for a count, and a sale in the middle of
counting is never called a loss.

## In the app

### Stock off the shelf in Review

When stock in the count's types is away, Review's summary says so, such as
"Off the shelf, not counted: 1 with the karigar". Under the product, it says
how many were away when it was counted, "not expected on the shelf".

```{r}
screenshot("review-off-shelf.png")
```

If a count finds more than expected while some are away, Review asks "did some
come back?" and offers **Back in**, so a piece that came back without being
recorded can be put right in the count.

After the count is completed, its Result and report list what was **off the
shelf** when each product was counted.

### Lines whose stock moved

Under a product counted before the counting ended, or recounted, Review says
what its expected stock was taken from, such as "Expected as of 11:42, when it was counted". A product whose count matches
none of the stock figures during its window shows "Stock moved while
counted", and waits under **Recount first** with its numbers hidden, like a
shortage (@sec-review). After the recount, the recount's own time becomes the
cut-off.

### Entries made after a count

Sometimes a sale is entered late, after the count that covered it was
completed, but dated before the product was counted. If that count had already
written the piece off as short, the shop would lose it twice: once in the
count, and again in the sale.

Jholok spots this when you record the movement. The sheet says, for example:
"This sale is dated before Showcase 2 counted Jhumka, which wrote 1 off as
short. Jholok will put 1 back so it isn't taken off twice." For a sale, it
asks for the **bill number**. Putting the piece back needs the owner, who
confirms with Face ID or the owner PIN. The correction is a [**cut-off
correction**](glossary.qmd#cut-off-correction) movement, logged with the sale.

If the owner doesn't confirm, the sale can still be saved without the
correction; the count then lists it as entered after completion, and its
write-off stays.

The count's Result and report list every entry made after completion, and the
next Review of those types lists the corrections, so nothing changes a past
count quietly.

## Why it works this way

- **What's away isn't lost.** Counting it as short would write off stock the
  shop still owns.
- **A shop doesn't close for a count.** The cut-off lets sales carry on during
  counting without spoiling it.
- **A late sale shouldn't count twice.** The cut-off correction puts back what
  the count took off, and only for that sale.
- **The owner decides corrections.** Changing what a completed count did is
  serious, so it needs the owner's proof, and it's logged and listed.

## How it works

For each product, the counting window runs from the first to the last time
its count changed, as the activity log records. Expected stock is the stock
in the shop at the window's end. If movements fell inside the window, Jholok
compares the count with the stock before each one and after it.

When every counted piece of a product had its tag read and the count has the
books (@sec-book-stock), expected comes from the tags themselves: the pieces
the books have on the shelf.

Completing a count freezes, for each product, its cut-off, how it was decided
and how much was off the shelf. Later movements never change what the count
compared.

## Try it

In the screenshot above, a `r tolower(out$name)` (`r out$variant`) went to
`r out$party` on `r date_words(out$sent_at)`, memo `r out$memo`. When the
chains were counted at `r open_count$location`, the shop owned
`r chain_owned` but had `r chain_in_shop` on the shelf. The counter counted
`r chain_line$counted`, so Review shows it
`r chain_line$counted - chain_in_shop` over, with
"`r out$units` with the karigar, not expected on the shelf".

One over, with one away: the likeliest story is that the chain came back from
the karigar and nobody recorded it. Tapping **Back in** records it as back
from when it was counted, and the line then matches.

**Question.** At 9:15 you count 6 pairs of hoops. At 9:20 a pair is sold. You
finish the earrings at 9:40. Is the hoops' count short?

::: {.callout-tip collapse="true"}
## Answer

No. The hoops were counted at 9:15, before the sale, when the shop had 6.
Their cut-off is 9:15, and your count matches the stock then. Had you
counted 5, they would be one short: at 9:15 the sale hadn't happened yet.
:::

::: {.remember}
- A count expects what's in the shop. Stock on jangad, with the karigar or
  kept for an order is listed apart, never as a loss.
- Each product's cut-off is when it was counted. Sales during counting are
  fine if the count matches the stock at some point.
- A product whose stock moved and matches nothing is recounted.
- A movement entered late, dated before a completed count, gets a cut-off
  correction that the owner confirms.
:::
