# Submit, review and complete {#sec-review}

```{r}
#| include: false
source(here::here("_common.R"))

open_count <- filter(shop$counts, is.na(completed_at))
scope_types <- unlist(open_count$scope)
in_scope <- shop$products |>
  filter(
    type %in% scope_types,
    created_at <= open_count$started_at,
    is.na(archived_at)
  )
lines <- shop$count_lines |>
  filter(count_id == open_count$count_id) |>
  rowwise() |>
  mutate(
    expected = stock_in_shop(shop, counted_at) |>
      filter(.data$sku == .env$sku) |>
      pull(in_shop)
  ) |>
  ungroup()
matched <- sum(lines$counted == lines$expected)
counted_units <- sum(lines$counted)

# The last earlier count of the same types: the whole-shop count, read over
# the necklaces only.
whole_shop <- filter(shop$counts, location == "Whole shop")
then_lines <- shop$count_lines |>
  filter(count_id == whole_shop$count_id) |>
  semi_join(filter(shop$products, type %in% scope_types), by = "sku") |>
  left_join(
    stock_on_hand(shop, whole_shop$started_at),
    by = "sku"
  )
then_matched <- sum(then_lines$counted == then_lines$on_hand)

# Showcase 1, completed: what its ticks did.
showcase_adjusted <- shop$movements |>
  filter(kind == "count_adjustment", grepl("Showcase 1", note)) |>
  left_join(select(shop$products, sku, name, variant, cost), by = "sku")
```

::: {.audience}
For approvers
:::

::: {.learn}
- How to open a count's Review, and who may.
- How to read Review, recount, and choose which differences to accept.
- What completing a count does, and how to sign it.
:::

## The idea

When the counters submit a count, it waits for an approver. The approver's
job is to look at what was found, check anything doubtful, and decide which
differences become changes to the stock. Until the count is completed,
nothing has changed.

Review is where the expected numbers appear for the first time, so only an
approver or the owner who didn't count in it may open it (@sec-setting-up).
The owner may approve a count they counted in only when nobody else can. It
starts with safe choices already made:

- **A shortage must be recounted** before it can be accepted. Its numbers stay
  hidden until then, and the recount is blind.
- **Overs start unticked.** An over usually means a missing record: enter that
  instead.
- **A product whose stock moved during counting** and matches nothing must be
  recounted too (@sec-cut-off).
- **Products nobody counted keep their stock**, unless the approver chooses
  Set to zero.

The approver can change the ticks, but can't accept a shortage that hasn't
been recounted. These rules are built into the app, not just the screen.

## In the app

### Open Review

Open the submitted count from the Count tab and tap **Review**. The app asks
for an approver's PIN. Only people who may approve are offered; those who
counted are named as unable to.

```{r}
screenshot("approve-to-review.png")
```

If the count was loaded with the books, and counting went on more than half an
hour after the time they're from, the app warns first (@sec-book-stock).

### Read Review

The card at the top says what was counted and how it compares:

- how many products were counted, and in what units;
- how many were **short** and **over**, in units and at cost;
- a sentence: how many matched, what happens to those not counted, and how
  this count compares with the last count of the same types.

```{r}
screenshot("review.png")
```

Below it come the sections:

- **Not identified**: box photo pieces nobody has identified. The count can't
  be completed until a counter identifies them (@sec-identify-later).
- **Recount first**: shortages, and products whose stock moved, with their
  numbers hidden.
- [**Discrepancies**](glossary.qmd#discrepancy): every other difference, each with its tick, the ledger
  and counted numbers, and the value at cost. The line at the end says what
  completing would change, such as "Net if completed: −₹1.29 L at cost."
- **Not counted**: products in the count that nobody counted, with the choice
  **Keep their stock** or **Set to zero**.
- **Matched**: folded away.

```{r}
screenshot("review-lower.png")
```

### Recount

Tap a product under **Recount first** to recount it. The recount is blind:
the first count and the ledger stay hidden until you tap **Done**. The app
keeps the first count and when the recount happened, even if the recount
confirms it.

```{r}
screenshot("review-recount-first.png")
```

```{r}
screenshot("recount.png")
```

You can recount any line under **Discrepancies** or **Not counted** by
tapping it.

### Choose, then complete

Tick the differences to accept. A ticked line sets stock to the count; an
unticked one leaves the stock as it is. Then tap **Complete count**, or **Complete and
adjust 3** when three lines will change.

```{r}
screenshot("complete-count.png")
```

The sheet says what will change, and what won't. Give a **reason**: it goes on
the count and on every adjustment it writes, as "Stock count · Showcase 1 ·
your reason". Recent reasons are offered to tap. Then **Sign with PIN and
complete**.

If the count needs more counting, close Review and tap **Reopen for counting**
on the count, with a reason. The log records that anything counted after
Review was counted after it.

## Why it works this way

- **A second person decides.** The approver didn't count (unless the owner is
  the only one who can approve, and the count then says so), so a loss can't be
  counted away and approved by the same person.
- **A loss someone counted is never accepted on one count.** The recount is cheap; a wrong
  write-off isn't.
- **Hidden until recounted.** A recount that can see the ledger tends to find
  the ledger's number.
- **The reason is written on every adjustment**, so the ledger says why each
  change happened.
- **The PIN is a signature.** The count records who approved it, and the log
  can't be edited afterwards.
- **Comparisons are honest.** Review says a match rate rose or fell only when
  the change is bigger than chance would give, with at least 10 products on
  each side. Otherwise it says "Too few products to call that a change."

## How it works

Completing is one change, saved at once: if anything fails, nothing changes
and the count stays open. Jholok freezes every line's expected and counted
numbers, cost, cut-off and stock off the shelf, so the count reads the same
forever. For each ticked line it writes a **count adjustment** movement for
the difference, signed by the approver, with the reason. It records who
approved, when and why, and logs the completion.

A completed count can't be edited. Mistakes found later are put right with
new movements, or a new count.

## Try it

The demo's open count, at `r open_count$location`, covers the
`r tolower(scope_types)`. Its Review card would say: counted
`r nrow(lines)` of `r nrow(in_scope)` products ·
`r describe_units(counted_units, "piece")`; `r matched_words(matched, nrow(lines))`;
`r nrow(in_scope) - nrow(lines)`
not counted keep their stock. The last count of the necklaces was the whole
shop count on `r date_words(whole_shop$completed_at)`:
`r then_matched` of `r nrow(then_lines)`. Both are every product matched, so
Review says "The same share as then."

At the earlier count at Showcase 1, the approver recounted the short jhumka,
found it still short, and ticked it. The hoops that were over, because of a
customer's repair piece, stayed unticked. The counter had also found one
single jhumka earring: Review showed it beside the pairs, as "+ 1 single",
and it changed no stock. Completing wrote one count
adjustment:

```{r}
#| label: tbl-showcase-adjustment
#| tbl-cap: "What completing Showcase 1 wrote in the ledger."
showcase_adjusted |>
  transmute(
    Product = paste(name, variant, sep = ", "),
    Change = sub("-", minus_sign, sprintf("%+d", quantity), fixed = TRUE),
    `At cost` = rupees(quantity * cost),
    Note = note
  ) |>
  table_guide()
```

**Question.** A product shows 2 short. You want to accept it, but its tick
won't turn on. Why?

::: {.callout-tip collapse="true"}
## Answer

It hasn't been recounted. Tap it under **Recount first**, count it again
blind, and tap Done. If it's still short, it moves to Discrepancies, already ticked.
:::

::: {.remember}
- Only an approver or the owner who didn't count opens Review and completes.
- Shortages, and products whose stock moved, are recounted blind first.
- Overs start unticked; products not counted keep their stock unless you
  choose Set to zero.
- Give a reason and sign with your PIN. Completing changes stock only for the
  ticked lines, and the count can't be edited afterwards.
:::
