# Results and the count report {#sec-results}

```{r}
#| include: false
source(here::here("_common.R"))

completed <- shop$counts |>
  filter(!is.na(completed_at)) |>
  arrange(completed_at)

# Each completed count, read as the Count tab's record lists it.
count_record <- shop$count_lines |>
  inner_join(
    select(completed, count_id, location, scope, started_at, completed_at),
    by = "count_id"
  ) |>
  left_join(select(shop$products, sku, type), by = "sku") |>
  left_join(shop$types, by = "type") |>
  rowwise() |>
  mutate(
    expected = stock_in_shop(shop, started_at) |>
      filter(.data$sku == .env$sku) |>
      pull(in_shop)
  ) |>
  ungroup() |>
  mutate(difference = counted - expected)

units_words <- \(differences, units) {
  totals <- vapply(split(abs(differences), units), sum, numeric(1))
  totals <- totals[totals > 0]
  if (length(totals) == 0) {
    return("–")
  }
  list_words(describe_units(totals, names(totals)))
}

covered_words <- \(types) {
  if (length(types) == 0) "Whole shop" else list_words(types)
}

record_table <- count_record |>
  summarise(
    Date = date_words(first(completed_at)),
    Covered = covered_words(scope[[1]]),
    Matched = paste(sum(difference == 0), "of", n()),
    Short = units_words(difference[difference < 0], unit[difference < 0]),
    Over = units_words(difference[difference > 0], unit[difference > 0]),
    .by = count_id
  ) |>
  select(-count_id)

showcase <- filter(count_record, location == "Showcase 1")
showcase_matched <- sum(showcase$difference == 0)
adjusted <- shop$movements |>
  filter(kind == "count_adjustment", grepl("Showcase 1", note))
left_as_was <- sum(showcase$difference != 0) - nrow(adjusted)
```

::: {.audience}
For the owner and approvers
:::

::: {.learn}
- What a completed count's Result tells you.
- What's in the count report, and how to share it.
- How the Count tab keeps the record of every count.
:::

## The idea

Once a count is completed, it becomes a record: what was counted, what
was found, what changed, who counted and who approved, and why. That record
can't be edited. Jholok shows it on screen as the count's **Result**, and can
print it as a **count report**, a PDF to keep with the shop's accounts or show
an auditor or insurer.

## In the app

### The Result

Open a completed count from the Count tab. The card at the top leads with
what was counted, then what was short and over, in units and at cost. The
sentence ends with what changed: for the demo's count at Showcase 1,
"`r showcase_matched` of `r nrow(showcase)` matched
(`r share(showcase_matched, nrow(showcase))`)… Completed
`r date_words(completed$completed_at[completed$location == "Showcase 1"])`.
Stock adjusted for `r nrow(adjusted)` product; `r left_as_was` left as it
was."

```{r}
screenshot("result.png")
```

Below it:

- when the count was completed, and what it covered;
- where expected stock came from (Jholok's ledger, or the books);
- who counted, and who approved and why;
- the count's part of the activity log, and whether it's **intact**.

If the phone hasn't been backed up since, a reminder says "This count is
only on this phone until it's backed up" (chapter 15).

Then the lists:

- **Stock adjusted** and **Left unchanged**;
- where they apply, the tags not found or found outside the books
  (@sec-book-stock), what was off the shelf (@sec-cut-off), single earrings
  found, and pieces not in the catalogue (@sec-identify-later);
- entries made after completion;
- products not counted, and the products that matched.

```{r}
screenshot("result-lower.png")
```

### Share the report

Tap the **Export** button at the top and choose:

- **Count report (PDF)**: the full report, for printing or keeping;
- **Lines and log (CSV)**: every line of the count, and its log, for a
  spreadsheet.

```{r}
screenshot("result-export.png")
```

The owner confirms each time, with Face ID or the owner PIN (until People is
set up, Face ID or the phone's passcode). The first time,
Jholok asks for the shop's name and address, to print at the top. The files
are made only for sharing: once the share sheet closes, Jholok deletes its
copies.

Settings → Export also has **Count results**: every line of every completed
count, in one spreadsheet. While a count is open, exports that would show its
stock wait until it's reviewed or discarded.

### The record of every count

The Count tab keeps score. Under **Last counted**, each type shows how long
since it was counted. Under **Record accuracy**, every completed count is
listed with how many products matched and what was short and over. With
fewer than 8 counts, it's a table; from 8, a chart shows each count's match
rate, with the range chance allows.

```{r}
screenshot("counts.png")
```

## What's in the report

The report is printed in black and white, on A4, so it copies and files well.

```{r}
report_page(
  1,
  alt = paste(
    "The report's first page: the shop's name, the count, its sign-off,",
    "totals by type and the discrepancies."
  ),
  caption = "Page 1: the shop, the count, the sign-off, totals, discrepancies."
)
```

- **The first page**: the shop's name and address; the count's place, types
  and times; what expected stock came from; the **sign-off** (who started,
  counted, submitted, reviewed and approved, and the reason); the log's code;
  totals **by type**; and the **discrepancies**, each with how it was counted.

```{r}
report_page(
  2,
  alt = "A page listing every line of the count, by type.",
  caption = "Every line, by type, with how each was counted."
)
```

- **Every line** of the count, grouped by type.
- Where they apply: the tags, what was off the shelf, customers' pieces,
  pieces not in the catalogue, and entries made after completion.

```{r}
report_page(
  3,
  alt = paste(
    "A page with a box photo, its pieces numbered, and a table of each",
    "piece's product and how it was identified."
  ),
  caption = "A page for each box photo, its pieces numbered."
)
```

- **A page for each box photo**, with its pieces numbered and a table of what
  each was and how it was identified, and each tag scan's tags.

Every page carries the report's [**fingerprint**](glossary.qmd#fingerprint) at the bottom: a code made
from everything in the report. Printing it again gives the same code, unless
a product was renamed, the shop's name or address changed, or entries were
made after the count was completed; then the report says so. Making a report
is logged, with its fingerprint.

The report gives quantities, not rupee values.

## Why it works this way

- **A completed count is a record, not a draft.** Its counts and numbers can't
  be edited, so a report printed next year shows the same count.
- **The fingerprint shows the paper is genuine.** Print the report again
  from the app and compare the codes: if the paper was altered, they won't
  match. (A renamed product or a late entry also changes the code, and the
  new report says so.)
- **Exports ask the owner every time**, because a report shows the shop's
  stock, and once a file leaves the phone, Jholok can't take it back.
- **Accuracy is shown honestly.** No percentage on fewer than 10 products, and
  no "better" or "worse" unless the difference is bigger than chance.

## How it works

Completing a count freezes its counts and numbers. Product names and SKUs are
printed as they are on the day (the report says so), and entries made after
completion are added. The fingerprint is a code worked out from everything in
the report except the time it was printed. The log's code on the report is the code
of the count's log as it was at completion: if anyone altered the log, the report would
say so.

## Try it

The demo shop has completed two counts. The Count tab's record would list
them like this:

```{r}
#| label: tbl-record
#| tbl-cap: "Jholok Jewellers' completed counts, as the Count tab lists them."
table_guide(record_table)
```

With two counts, the record is a table: too few to draw a trend.

**Question.** An insurer asks for proof of your last stock count. What do you
send, and how can they tell it hasn't been changed?

::: {.callout-tip collapse="true"}
## Answer

The count's report, from Export → Count report (PDF). Every page carries its
fingerprint: printing the report again from the app gives the same code, as
long as nothing was renamed or entered late, and the report says when
something was.
:::

::: {.remember}
- A completed count's numbers can't change.
- The report shows the sign-off, totals, discrepancies, every line and every
  box photo, with a fingerprint on every page.
- Every export asks the owner first, and Jholok deletes its copy once it's
  shared.
- The Count tab keeps the record of every count.
:::
