# Keeping your data safe {#sec-keeping-data-safe}

```{r}
#| include: false
source(here::here("_common.R"))
```

::: {.audience}
For the owner
:::

::: {.learn}
- Where Jholok keeps the shop's data, and what a lost phone would cost.
- How to back up, and how to restore or move to a new phone.
- How Jholok keeps the data private, and what it does when something goes
  wrong.
:::

## The idea

Jholok keeps everything on the phone: the catalogue, every movement, every
count, the activity log, product photos and box photos. Nothing is sent to a
server or synced to the cloud. That keeps the shop's stock and costs private,
but it means **the phone holds the only copy until you make a backup**.

A **backup** is one file with everything in it. Saved somewhere other than
the phone (iCloud Drive, a computer), it lets you get the shop back if the
phone is lost, broken or replaced.

Jholok doesn't back up by itself. Make it a habit: after every completed
count, and at least once a week.

## In the app

### Back up

Open **Settings**, then **Backup**.

```{r}
screenshot(
  "settings-backup.png",
  caption = "Backup, in the demo shop, where backups are turned off."
)
```

The top says when you last backed up, and some of what's been added since
("Since then: 1 count completed and 14 box photos added"). A lost phone would
lose those, and every sale, receipt and other entry made since too. **A backup holds** lists what goes in: products, counts,
any count in progress, and box photos.

1. Tap **Back up now**.
2. Confirm it's you, with Face ID or the owner PIN (until People is set up,
   Face ID or the phone's passcode).
3. Choose where to save the file, such as **iCloud Drive** in the Files list,
   or a folder that a computer collects. The file is called something like
   "Jholok backup `r format(screens_taken_on)` 1730.jholokbackup".

The backup holds every product and photo, every count and movement, the
activity log, the box photos, and the preferences (currency, appearance,
lead time, book stock words and so on).

After a count is completed, its Result reminds you if it isn't backed up yet:
"This count is only on this phone until it's backed up." (@sec-results)

### Restore, or move to a new phone

To restore, open **Settings** → **Backup**, tap **Restore from a backup…**,
confirm it's you, and choose the backup file.

Before anything changes, Jholok shows the phone and the backup side by side:
when the backup was made, and how many products, counts, counts in progress
and box photos each has.
It warns that everything on the phone will be replaced. If the phone has data,
**Back up this phone first** saves it before you go on. Then tap **Replace
with backup**.

To move to a new phone, back up on the old one, install Jholok on the new
one, and restore the backup there. The lock setting and the shop's name and
address (Settings → Shop) aren't in the backup: set them again on the new
phone (@sec-setting-up).

### Privacy

- **The lock**: Jholok asks for Face ID or the phone's passcode when it opens
  and, by default, after a minute away; Settings → Security changes it
  (@sec-setting-up).
- **The cover**: whenever Jholok isn't in front, a plain cover hides it from
  the list of open apps.
- **The owner, every time**: every export, backup, restore, report and box
  photo export asks the owner to confirm, with Face ID or the owner PIN (or the phone's
  passcode, until People is set up).
- **Nothing left behind**: files made for sharing are deleted as soon as the
  share sheet closes, and any left over are deleted when Jholok opens.

### Box photos and space

Box photos take space. **Settings → Photo counting → Storage** says how much,
and can **remove photos of counts completed over 12 months ago**. Each box's
pieces and numbers stay; only the photos go.

## Why it works this way

- **On the phone, not in the cloud**, so the shop's stock and costs are seen
  only by the people who hold the phone.
- **One file**, so a backup is easy to save, copy and keep.
- **Side by side before restoring**, so nobody replaces new data with an old
  backup by mistake.
- **The log remembers restores.** After a restore, the log records it and
  where the log it replaced had got to. Going back to an older backup always
  shows.
- **Exports ask every time**, because once a file leaves the phone, Jholok
  can't take it back.

## How it works

### Saving never fails silently

Every change is saved at once. If a save fails (most often because the phone
is nearly full), Jholok puts everything back as it was, so the screen matches
what's stored, and says so: "The send out wasn't saved, so it's back as it
was. If the phone is nearly full, free some space and try again." A count
whose completion fails stays open, with no stock changed.

### If Jholok can't open its data

If the data can't be opened (after a failed update, say), Jholok doesn't
stop. It shows "Jholok can't open its data", and says nothing has been
deleted. You can **Try again**, **Restore from a backup…**, or **Save the data
file…**, which saves the raw data so it can be repaired. That file isn't a
backup and can't be restored.

### Restoring safely

Jholok first copies the backup next to your data, checks it and brings it up
to date with the app. Only then does it replace your data. If the
phone switched off part way, Jholok finishes the job the next time it opens.
A backup made by an older version of Jholok is brought up to date as it's
restored. One made by a newer version can't be restored until you update the
app.

### Diagnostics

**Settings → Diagnostics**, under About, keeps reports of crashes and hangs from iOS, and a
log of what went wrong in the app, such as a failed save or a camera that
stopped. It never records stock, products, prices or tags. Nothing leaves the
phone unless the owner taps **Share diagnostics**, which makes one file to
send to whoever supports the app.

## Try it

**Question.** The shop's phone is stolen. What do you need to get the shop
back, and what would be lost?

::: {.callout-tip collapse="true"}
## Answer

The latest backup file, saved off the phone. Install Jholok on a new phone
and restore it. Everything entered since that backup would be lost: the counts, products and
box photos in the Backup screen's "Since then" line, and every sale, receipt
and other entry too. That's why it's worth backing up after every count.
:::

::: {.remember}
- The phone holds the only copy until you back up. Jholok doesn't back up by
  itself.
- Back up after every completed count, and at least weekly, to iCloud Drive
  or a computer.
- Restoring shows the phone and the backup side by side first, and replaces
  everything.
- Every export and backup asks the owner, and shared files are deleted
  afterwards.
:::
