# Stories from the shop {#sec-stories}

```{r}
#| include: false
source(here::here("_common.R"))

demo <- load_shop(counting = FALSE)
names_of <- select(demo$products, sku, name, variant, cost)

# 1. The first whole-shop count.
whole_shop <- filter(demo$counts, location == "Whole shop")
first_count <- demo$count_lines |>
  filter(count_id == whole_shop$count_id) |>
  left_join(
    stock_in_shop(demo, whole_shop$started_at) |>
      rename(expected = in_shop),
    by = "sku"
  ) |>
  left_join(names_of, by = "sku") |>
  mutate(difference = counted - expected)
first_differences <- filter(first_count, difference != 0)

# 2. Dhanteras: the busiest day for gold coins.
dhanteras <- festivals |>
  filter(festival == "Dhanteras", date < as.Date("2026-01-01")) |>
  mutate(date = date + demo$shift_days)
coins_that_day <- demo$movements |>
  filter(
    kind == "sold",
    as.Date(date, tz = shop_time_zone) == dhanteras$date,
    startsWith(sku, "COI")
  ) |>
  summarise(sold = -sum(quantity), .by = sku) |>
  left_join(names_of, by = "sku") |>
  arrange(desc(sold))

# 3. The pendant below zero.
pendant <- stock_history(demo, "NEC-0003")
pendant_now <- sum(pendant$quantity)
early_sale <- slice_tail(pendant, n = 1)

# 4. The karigar's chain.
karigar <- demo$out_records |>
  filter(place == "karigar") |>
  left_join(names_of, by = "sku")
days_late <- as.integer(
  difftime(demo$end_at, karigar$due_back, units = "days")
)

# 5. Showcase 1.
showcase <- filter(demo$counts, location == "Showcase 1")
```

::: {.audience}
For everyone
:::

::: {.learn}
- How the ideas and screens in this guide fit together in a shop's year.
- What to do in situations every jeweller meets.
:::

These are moments from Jholok Jewellers' year. The figures are the demo
shop's own, so you can open the app's demo and find each of them.

## The first full count

Bidisha decided to count the whole shop at the end of the financial year,
before opening on `r date_words(whole_shop$started_at, "%B")`. She had set
up People the week before (@sec-setting-up): Ramesh and Meena to count, Sita
to approve.

Ramesh started the count in the Count tab, with **All types**, and he and
Meena counted tray by tray, each signed in with their own PIN
(@sec-a-count). Neither could see what the ledger expected. They finished in
two hours and Ramesh submitted.

Sita opened Review. `r matched_words(sum(first_count$difference == 0), nrow(first_count))`.
These didn't:

```{r}
#| label: tbl-whole-shop-count
#| tbl-cap: "What the first whole-shop count found."
first_differences |>
  transmute(
    Product = paste(name, variant, sep = ", "),
    Expected = expected,
    Counted = counted,
    `At cost` = rupees(difference * cost)
  ) |>
  table_guide()
```

The shortages waited under **Recount first**, so Sita recounted them herself,
blind. They were still short, and she found no missing sale to explain them.
She recounted the coins too, checked that no delivery of coins was missing,
and accepted the extra one. She ticked all three and completed the count with
a reason, signing with her PIN (@sec-review). Each accepted difference became
a count adjustment in the ledger.

**The lesson:** the first count sets the baseline. From then on, every count
is compared with the last one of the same types.

## Dhanteras

On Dhanteras, `r date_words(dhanteras$date, "%B")`, the shop sold
`r sum(coins_that_day$sold)` coins in one day:

```{r}
#| label: tbl-dhanteras
#| tbl-cap: "Coins sold on Dhanteras."
coins_that_day |>
  transmute(Coin = paste(name, variant, sep = ", "), Sold = sold) |>
  table_guide()
```

Every sale was recorded as it happened, so the ledger kept up. The Overview's
sales chart labelled that week with its festival (@sec-overview), and **Needs
attention** lists any product that, at its new pace, won't last the supplier
lead time.

**The lesson:** after a rush, count what moved fastest. Busy days are when a
sale is most easily missed.

## The pendant below zero

The Overview showed one product **Below zero**: the Lakshmi pendant, at
`r sub("-", minus_sign, pendant_now, fixed = TRUE)`. Its last movement was a sale on
`r date_words(early_sale$date)`, with the note "`r early_sale$note`".

The pendant had sold out, a delivery of 6 had arrived, and a customer bought
one before anyone entered the delivery. Bidisha opened the pendant and tapped
**Receive**, entered 6, and set the date to the day the delivery came. The
stock became `r pendant_now + 6`, and the warning went away
(@sec-movements).

**The lesson:** below zero always means a missing movement. Find it, enter it
with its real date, and the stock is right again.

## The customer's repair piece

At the count at `r showcase$location` on
`r date_words(showcase$completed_at, "%B")`, the hammered hoops came out
one pair over. Sita, approving, didn't accept it straight away: an over
usually means a missing record. Ramesh, who led the count, told her why: a
customer's repair piece had been left in the tray.

The extra pair wasn't the shop's stock at all, so Sita left the line
unticked. The hoops' stock didn't change, and the Result lists the line under
**Left unchanged** (@sec-results).

**The lesson:** keep customers' pieces out of the trays, or mark them, and
never make stock up to a count without asking why.

## The karigar's chain

A `r tolower(karigar$name)` (`r karigar$variant`) had gone to
`r karigar$party` for repair, memo `r karigar$memo`, due back in 30 days. By
then it was `r days_late` days late, and the Overview's **Needs attention**
listed it: "1 memo more than a week overdue" (@sec-movements).

Bidisha called the karigar. The chain had come back the week before and was
already in the showcase, but nobody had recorded it. Sita opened the chain,
tapped **⋯** beside the memo under **Off the shelf**, chose **Back in**, set
**Back on** to the day it came back, and signed with her PIN. The chain was
in the shop again, and the next count expected it on the shelf
(@sec-cut-off).

**The lesson:** record pieces back in when they come back. Otherwise the next
count finds them "over", and the overdue list keeps nagging.

## A new member of staff

When a new assistant joined, Bidisha added them in Settings → People as a
**Counter**, and let them choose their own PIN (@sec-setting-up). Their first
count was a small one, the rings in one showcase, with Ramesh beside them to
show how + and the number pad work, and how − and Undo put a slip right.

Counting is blind, so the ledger couldn't sway the new counter. And a
counter can never approve a count. The activity log shows each figure they entered.

**The lesson:** start new staff on small counts. The app's rules protect the
shop and them: nobody can be blamed for a figure someone else entered.

::: {.remember}
- Count the whole shop once to set a baseline; then count often, a few types
  at a time.
- Record every movement when it happens, with its real date.
- A difference is a question: below zero, over and short each have a usual
  cause worth finding.
- Record pieces back in when they come back from jangad or the karigar.
:::
